### 2013-01-14 documents/thesis/TQLBO-paper/Paper-Siggraph/paper.tex
msg

%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,review]{acmsiggraph}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amssymb}
\usepackage{graphicx}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
% for the following cases use the listed document class option:
% [annual] - Technical paper accepted for presentation at the ACM SIGGRAPH 
%   or SIGGRAPH Asia annual conference.
% [sponsored] - Short or full-length technical paper accepted for 
%   presentation at an event sponsored by ACM SIGGRAPH
%   (but not the annual conference Technical Papers program).
% [abstract] - A one-page abstract of your accepted content
%   (Technical Sketches, Posters, Emerging Technologies, etc.). 
%   Content greater than one page in length should use the "[sponsored]"
%   parameter.
% [preprint] - A preprint version of your final content.
% [review] - A technical paper submitted for review. Includes line
%   numbers and anonymization of author and affiliation information.

% When you submit your paper for review, please use the \TOGonlineID''
% command to include the online ID value assigned to your paper by the
% submission management system. Replace '45678' with the value you were
% assigned.
\TOGonlineid{XXXXX}

% If you are preparing a preprint of your accepted paper, and your paper
% will be published in an issue of the ACM "Transactions on Graphics''
% journal, replace the "0'' values in the commands below with the correct
% volume and number values for that issue - you'll get them before your
% final paper is due.
\TOGvolume{0}
\TOGnumber{0}

% The TOGarticleDOI' command accepts the DOI information provided to you
% during production, and which makes up the URLs which identifies the ACM
% article page and direct PDF link in the ACM Digital Library.
% Replace "1111111.2222222'' with the values you are given.
\TOGarticleDOI{1111111.2222222}

% If you would like to include links to personal repositories for auxiliary
% material related your research contribution, you may use one or more of
% these commands to define an appropriate URL. The "\TOGlinkslist'' command
% found just before the first section of your document will add hyperlinked
% icons to your document, in addition to hyperlinked icons which point to
% the ACM Digital Library article page and the ACM Digital Library-held PDF.
\TOGprojectURL{}
\TOGvideoURL{}
\TOGdataURL{}
\TOGcodeURL{}

% Paper title.
\title{Curvature Enhanced Shape With and Adapted Laplacian Operator For Hybrid Quad/Triangle Meshes}

% Author and Affiliation (single author).
%\author{Name \thanks{e-mail: name@unknown.uu}\\ Research Institute}

% Author and Affiliation (multiple authors).
\author{Alexander Pinzon\thanks{e-mail: apinzonf@gmail.com}
\qquad  Eduardo Romero\thanks{e-mail:edromero@unal.edu.co}
 \\Cimalab Research Group\\National University of Colombia}
%\and Ton Roosendaal\thanks{e-mail:ton@blender.org}\\ Blender CEO}

% The ``pdfauthor'' command accepts the authors of the work,
% comma-delimited, and adds this information to the PDF metadata.
%\pdfauthor{Alexander Pinzon, Eduardo Romero, Ton Roosendaal}
\pdfauthor{Anonymous}

% Keywords that describe your work.
\keywords{laplacian smooth, curvature, sculpting, subdivision surface}

\makeatother

\usepackage{babel}
\begin{document}

\teaser{\includegraphics[width=1\textwidth]{images/spectrum} \caption{\label{fig:Spectrum}A set of 48 successive curvature enhanced shapes,
from $\lambda=0.0$ in blue to $\lambda=-240.0$ in red, with steps
of $-5.0$.}
}

\maketitle
\maketitle 
\begin{abstract}
This paper proposes a novel modeling method for constructing a hybrid quad/triangle
mesh by enhancing the global object curvature. This method uses an
adapted extension of the Laplace Beltrami operator, resulting in a large variety of meshes. Along with the
method, this work presents new applications in sculpting and
modeling, with subdivision of surfaces and weight vertex groups. A
series of graphics examples demonstrates the quality, predictability
and flexibility of the method in a real production environment with
software blender.
\end{abstract}

\begin{CRcatlist}
\CRcat{I.3.5}{Computer Graphics}{Computational Geometry and Object Modeling
}{Modeling packages}
\end{CRcatlist}
\keywordlist

\TOGlinkslist

%\copyrightspace

\section{Introduction}

Over the last years several, modeling techniques able to generate
a variety of realistic shapes, have been developed \cite{Botsch2006}.
Editing techniques have evolved from affine transformations to advanced
tools like sculpting \cite{Coquillart1990,Galyean1991,Stanculescu2011},
editing, creation from sketches \cite{Igarashi1999,Gonen2012}, and
complex interpolation techniques \cite{Sorkine2004,Zhou2005}. Catmull-Clark
based methods however require to interact with a minimun number of
control points for any operation to be efficient, or in other words,
a unicity condition is introduced by demanding a smooth surface after
any of these shape operations. Hence, traditional modeling methods
for subdividing surfaces from coarse geometry have become widely popular
\cite{Catmull1978,Stam1998}. These works have generalized a uniform
B-cubic spline knot insertion to meshes, some of them adding some
type of control, for instance with the use of creases to produce sharp
edges \cite{DeRose1998}, or the modification of some vertex weights
to locally control the zone of influence \cite{Biermann2000}. Nevertheless,
these methods are difficult to deal with since they require a large
number of parameters and a very tedious customization. Instead, the
presented method requires a single parameter that controls the global
curvature, which is used to maintain realistic shapes, creating a family of different versions of the same
object and therefore preserving the detail of the original model and
a realistic appearance. Interest in meshes composed of triangles and quads has lately increased because of the flexibility of modeling tools such as Blender 3D \cite{blender}. Nowadays, many artists use a manual connection of a couple of vertices to perform animation processes and interpolation \cite{Mullen2007}. It is then of paramount importance to develop operators that easily interact with such meshes, eliminating the need of preprocessing the mesh to convert it to triangles. The curvature enhancement and shape exaggeration can thus be used as such operator in the sculpting process, when inflating a shape since current brushes end up by losing detail when moving vertices \cite{Stanculescu2011}. In contrast, the presented method inflates a mesh by moving the vertices towards the reverse curvature direction, conserving the shape and sharp features of the model.

\textbf{Contributions} This work presents an extension of the Laplace
Beltrami operator for hybrid quad/triangle meshes, representing a
larger mesh spectrum from what has been presented so far. The method
eliminates the need of preprocessing and allows preservation of the
original topology. Likewise, along with this operator, it is proposed
a method to generate a family of parameterized shapes, in a robust
and predictable way. This method enables customization of the smoothness
and curvature, obtained during the subdivision surfaces process. Finally,
it is proposed a new brush for enhancing the silhouette mesh features
in modeling and sculpting.

This work is organized as follows: Section \ref{sub:1.1-Related-work}
presents works related to the Laplacian mesh processing, digital sculpting,
and offsetting methods for polygonal meshes; In section \ref{sec:Laplacian-Smooth}
, it is described the theoretical framework of the Laplacian operator
for polygon meshes; In section \ref{sec:Proposed-Method}, it is presented
the method for curvature enhancement and applications of subdivision
of surfaces and sculpting; finally some Laplacian operator results,
to hybrid quad/triangle meshes are graphically shown as well as results
of the curvature enhancement applications in sculpting, subdivision
and modeling.

\section{Related work\label{sub:1.1-Related-work}}

Many tools have been developed for modeling, based on the Laplacian
mesh processing. Thanks to the advantages of the Laplacian operator,
these different tools preserve the surface geometric details when
using them for different processes such as free-form deformation,
fusion, morphing and other application\cite{Sorkine2004}.

Offset methods for polygon meshing, based on the curvature defined
by the Laplace Beltrami operator, have been developed. These methods
adjust the shape offset by a constant distance, with enough precision
as to minimize the Hausdorff error. Nevertheless, these methods fail
to conserve sufficient detail because of the smoothing, a crucial
issue which depends on the offset size \cite{Zhuo2012}. In volumetric
approaches, in case of point-based representations, the offset boundary
computation is based on the distance field and therefore when calculating
such offset, the topology of the model may be different to the original
\cite{Chen2011}.

Gai et al. \cite{Gal2009} propose automatic feature detection and shape edition
with feature inter-relationship preservation. They define salient
surface features like ridges and valleys, characerized by their first
and second order curvature derivatives \cite{Ohtake2004}, and angle-based
threshold. Likewise, curves have been also classified as planar or
non-planar, approximated by lines, circles, ellipses and other complex
shapes. In such case, the user defines an initial change over several
features which is propagated towards other features, based
on the classified shapes and the inter-relationships between them.
This method works well with objects that have sharp edges, composed
of basic geometric shapes such as lines, circles or ellipses. However,
the method is very limited when models are smooth since
it cannot find the proper features to edit.

Digital sculpting have been traditionally approached either under
a polygonal representation or a voxel grid-based method. Brushes for
inflation operations only depend on the vertex normal \cite{Stanculescu2011}.
In grid-based sculpting, some other operations have allowed to add
or remove voxels since production of polygonal meshes require a processing of
isosurfaces from a volume \cite{Galyean1991}. The drawback comes from the difficulty of maintaining the surface details
during larger scale deformations.

\section{Laplacian Smooth\label{sec:Laplacian-Smooth}}

\begin{figure*}[t]
\includegraphics[width=1\textwidth]{images/beltrami}

\caption{\label{fig:LBO-basic-5-TQ}The 5 basic triangle-quad cases with common
vertex $V_{i}$ and the relationship with $V_{j}$ and $V_{j}^{\prime}$.
(a) Two triangles {[}Desbrun 1999{]}. (b) (c) Two quads and one quad
{[}Xiong 2011{]}. (d) (e) Triangles and quads (TQLBO).}
\end{figure*}

Computer objects, reconstructed from the real
world, are usually noisy. Laplacian Smooth techniques allow a proper
noise reduction on the mesh surface with minimal shape changes,
while still preserving a desirable geometry as well as the original shape.

Many smoothing Laplacian functionals regularize the surface energy by controling the total surface curvature $S$.

\begin{equation}
E\left(S\right)=\int_{S}\kappa_{1}^{2}+\kappa_{2}^{2}dS\label{eq:total_cuarvature_integral}
\end{equation}

Where $\kappa_{1}$ and $\kappa_{2}$ are the two principal curvatures
of the surface $S$.

\subsection{Gradient of Voronoi Area}

\begin{figure}[h]
\includegraphics[width=1\columnwidth]{images/voronoi_region}

\caption{\label{fig:voronoi_region}Area of the Voronoi region around $v_{i}$
in dark blue.\protect \\
 $v_{j}$ belong to the first neighborhood around $v_{i}$. $\alpha_{j}$ and $\beta_{j}$ are the
opposite angles to the edge $\protect\overrightarrow{v_{j}-v_{i}}$. }
\end{figure}

Consider a surface $S$, composed of a set of triangles, around the vertex
$v_{i}$. Let us define the \textit{Voronoi region} of $v_{i}$, as
show in figure \ref{fig:voronoi_region}, The area change produced
by the movement of $v_{i}$ is called the gradient of the \textit{Voronoi region
\cite{Pinkall1993,Desbrun1999,Meyer2003}.}

\begin{equation}
\nabla A=\frac{1}{2}\underset{j}{\sum}\left(\cot\alpha_{j}+\cot\beta_{j}\right)\left(v_{i}-v_{j}\right)\label{eq:eq_gradient_voronoi_area}
\end{equation}

If the gradient in equation (\ref{eq:eq_gradient_voronoi_area}) is normalized 
by the total area of the first neighborhood around $v_{i}$, the \textit{discrete
mean curvature normal} of a surface $S$  is obtained, as shown in equation (\ref{eq:discrete_mean_curvature_normal}).

\begin{equation}
2\kappa\mathbf{n}=\frac{\nabla A}{A}\label{eq:discrete_mean_curvature_normal}
\end{equation}

\subsection{Laplace Beltrami Operator}

The \textit{Laplace Beltrami operator} LBO, noted as $\triangle_{g}$,
is used for measuring the mean curvature normal to the Surface $S$ \cite{Pinkall1993}.

\begin{equation}
\triangle_{g}S=2\kappa\mathbf{n}\label{eq:def_LBO}
\end{equation}

The LBO has desirable properties: the LBO points out to the direction
of the minimal surface area, minimizing the energy of the total surface curvature $S$ at equation (\ref{eq:total_cuarvature_integral}).

\section{Proposed Method\label{sec:Proposed-Method}}

The method allows edition of geometric features using the curvature
enhancement and smoothing, generating a parameterized family of shapes
from a set of vertices which represent a coarse sketch of the desired
model. Our approach can be mixed with traditional or uniform subdivision
surfaces methods, is iterative and converges towards a continuous
and smooth version of the original model.

Our method allows the use of soft constraints weighting the effect
of smoothing at each vertex based on a normalized weight, which are
assigned to the control vertices of the original mesh or. The weights
of the new vertices resulting from the subdivisions are calculated
by interpolation, allowing modifying the behavior of the method on
specific regions of the original model.

Our work presents an extension of the Laplace Beltrami operator for
hybrid quad/triangle meshes, allows using mixed arbitrary types of
mesh representation, exploiting the basic geometrical relationships
facilitating and ensuring convergence of the algorithm and similar
shapes consistent with the original shape against the other methods.

\subsection{Laplace Beltrami Operator for Hybrid Quad/Triangle Meshes TQLBO\label{sub:Laplace-Beltrami-operator}}

Given a mesh $M=\left(V,Q,T\right)$, with vertices $V$, quads $Q$,
triangles $T$.

The area of $1$-ring neighborhood ($N_{1}$) with shared face quad
or triangle to vertex $v_{i}$ in $M$ is.

\begin{center}
$A\left(v_{i}\right)=A\left(Q_{N_{1}\left(v_{i}\right)}\right)+A\left(T_{N_{1}\left(v_{i}\right)}\right)$.
\par\end{center}

\begin{figure}[h]
\includegraphics[width=1\columnwidth]{images/quad_xiong}

\caption{\label{fig:quad_xiong}$t_{j1}^{*}\equiv\vartriangle v_{i}v_{j}v_{j}^{\prime},\, t_{j2}^{*}\equiv\vartriangle v_{i}v_{j}^{\prime}v_{j+1},\, t_{j3}^{*}\equiv\vartriangle v_{i}v_{j}v_{j+1}$
Triangulations of the quad with common vertex $v_{i}$ proposed by
{[}Xiong 2011{]} to define Mean LBO.}
\end{figure}

Applying the mean average area according to \cite{Xiong2011} of all
possible triangulations for each quad to $A\left(Q_{N_{1}\left(v_{i}\right)}\right)$
as show in figure \ref{fig:quad_xiong}.

\begin{center}
$A\left(v_{i}\right)=\frac{1}{2^{m}}\overset{m}{\underset{j=1}{\sum}}2^{m-1}A\left(q_{j}\right)+\overset{r}{\underset{k=1}{\sum}}A\left(t_{k}\right)$
\par\end{center}

Where $q_{1},q_{2},...,q_{j},...,q_{m}\in Q_{N_{1}\left(v_{i}\right)}$
and $t_{1},t_{2},...,t_{k},...,t_{r}\in T_{N_{1}\left(v_{i}\right)}$.

\begin{equation}
A\left(v_{i}\right)=\frac{1}{2}\overset{m}{\underset{j=1}{\sum}}\left[A\left(t_{j1}^{*}\right)+A\left(t_{j2}^{*}\right)+A\left(t_{j3}^{*}\right)\right]+\overset{r}{\underset{k=1}{\sum}}A\left(t_{k}\right)\label{eq:area_1_ring_triangles_quads}
\end{equation}

Applying the gradient operator to (\ref{eq:area_1_ring_triangles_quads}).

\begin{equation}
\nabla A\left(v_{i}\right)=\frac{1}{2}\overset{m}{\underset{j=1}{\sum}}\left[\nabla A\left(t_{j1}^{*}\right)+\nabla A\left(t_{j2}^{*}\right)+\nabla A\left(t_{j3}^{*}\right)\right]+\overset{r}{\underset{k=1}{\sum}}\nabla A\left(t_{k}\right)\label{eq:EqGrad}
\end{equation}

According to (\ref{eq:eq_gradient_voronoi_area}), we have.

\begin{center}
$\nabla A\left(t_{j1}^{*}\right)=\frac{\cot\theta_{j3}\left(v_{j}-v_{i}\right)+\cot\theta_{j2}\left(v_{j}^{\prime}-v_{i}\right)}{2}$
\par\end{center}

\begin{center}
$\nabla A\left(t_{j2}^{*}\right)=\frac{\cot\theta_{j5}\left(v_{j}^{\prime}-v_{i}\right)+\cot\theta_{j4}\left(v_{j+1}-v_{i}\right)}{2}$
\par\end{center}

\begin{center}
$\nabla A\left(t_{j3}^{*}\right)=\frac{\cot\theta_{j6}\left(v_{j}-v_{i}\right)+\cot\theta_{j1}\left(v_{j+1}-v_{i}\right)}{2}$
\par\end{center}

\begin{center}
$\nabla A\left(t_{k}\right)=\frac{\cot\alpha_{k}\left(v_{k}-v_{i}\right)+\cot\beta_{k+1}\left(v_{k+1}-v_{i}\right)}{2}$
\par\end{center}

Thus triangles and quads configurations of the 1-ring neighborhood
faces adjacent to $v_{i}$ can be simplified to five cases shown in
figure \ref{fig:LBO-basic-5-TQ}.

Then according to equation (\ref{eq:discrete_mean_curvature_normal}),
(\ref{eq:def_LBO}), and five simple cases defined in figure \ref{fig:LBO-basic-5-TQ}
the TQLBO (Triangle-Quad LBO) of $v_{i}$ is.

\begin{equation}
\triangle_{g}\left(v_{i}\right)=2\kappa\mathbf{n}=\frac{\nabla A}{A}=\underset{v_{j}\in N_{1}\left(v_{i}\right)}{\frac{1}{2A}\sum}w_{ij}\left(v_{j}-v_{i}\right)
\end{equation}

\begin{equation}
w_{ij}=\begin{cases}
\left(\cot\alpha_{j}+\cot\beta_{j}\right) & \mbox{case }\mathit{a.}\\
\frac{1}{2}\left(\cot\theta_{\left(j-1\right)1}+\cot\theta_{\left(j-1\right)4}+\cot\theta_{j3}+\cot\theta_{j6}\right) & \mbox{case \ensuremath{\mathit{b}}.}\\
\left(\cot\theta_{j2}+\cot\theta_{j5}\right) & \mbox{case \ensuremath{\mathit{c}}.}\\
\frac{1}{2}\left(\cot\theta_{j3}+\cot\theta_{j6}\right)+\cot\beta_{j} & \mbox{case \ensuremath{\mathit{d}}.}\\
\frac{1}{2}\left(\cot\theta_{\left(j-1\right)1}+\cot\theta_{\left(j-1\right)4}\right)+\cot\alpha_{j} & \mbox{case \ensuremath{\mathit{e}}.}
\end{cases}\label{eq:TQLBO_wij}
\end{equation}

We define a Laplacian operator as a matrix equation

\begin{equation}
L\left(i,j\right)=\begin{cases}
-\frac{1}{2A_{i}}w_{ij} & \mbox{if }j\in N\left(v_{i}\right)\\
\frac{1}{2A_{i}}\underset{j\in N\left(v_{i}\right)}{\sum}w_{ij} & \mbox{if }i=j\\
0 & \mbox{otherwise}
\end{cases}\label{eq:TQLBO_Simple_Matrix}
\end{equation}

Where $L$ is a $n\times n$ matrix, $n$ is the number of vertices
of a given mesh $M$, $w_{ij}$ is the TQLBO defined in equation (\ref{eq:TQLBO_wij}),
$N\left(v_{i}\right)$ is the 1-ring neighborhood with shared face
to $v_{i}$, $A_{i}$ is the ring area around $v_{i}$.

Normalized version of the TQLBO as a matrix equation

\begin{equation}
L\left(i,j\right)=\begin{cases}
-\frac{w_{ij}}{\underset{j\in N\left(v_{i}\right)}{\sum}w_{ij}} & \mbox{if }j\in N\left(v_{i}\right)\\
\delta_{ij} & \mbox{otherwise}
\end{cases}\label{eq:TQLBO-Normalized_Matrix}
\end{equation}

Where $\delta_{ij}$ being the Kronecker delta function.

\subsection{Curvature Enhancing\label{sub:Curvature-Enhancing}}

\begin{figure*}[t]
\includegraphics[width=1\textwidth]{images/teaser_cup}

\caption{\label{fig:Subdivision-Cups} Family of cups generated with our method
from coarse model to enhancing the curvature obtained from Catmull-Clark
Subdivision and the use of constraints over coarse model with weight
vertex group in red.}

\end{figure*}

The curvature enhancing uses the inverse direction of the curvature
flow to move the vertices in the portions of the mesh with most curvature.
In this equation we use a diffusion process:

\begin{center}
$\frac{\partial V}{\partial t}=\lambda L\left(V\right)$ 
\par\end{center}

To solve the equation above we use implicit integration and a normalized
version of TQLBO matrix.

\begin{equation}
\left(I-\left|\lambda dt\right|W_{p}L\right)V^{\prime}=V^{t}\label{eq:Lineal_System_with_wp}
\end{equation}

\begin{center}
$V^{t+1}=V^{t}+\mbox{sign}\left(\lambda\right)\left(V^{\prime}-V^{t}\right)$
\par\end{center}

The vertices $V^{t+1}$are enhanced along their inverse curvature
normal directions by solving the linear system: $Ax=b$, where $A=I-\left|\lambda dt\right|W_{p}L$,
$L$ is the Normalized TQLBO defined in the equation (\ref{eq:TQLBO-Normalized_Matrix}),
$x=V^{\prime}$ are the smoothing vertices, $b=V^{t}$ are the actual
vertices positions, $W_{p}$ is a diagonal matrix with weight vertex
group, and $\lambda dt$ is the enhance factor that support negative
and positive values, negative for enhancing positive for smoothing.

The curvature cannot be calculted at the border of the meshes that
are not closed, for that reason we use the scale-dependent operator
proposed by \cite{Desbrun1999}.

Our method was designed for use with weighted vertex groups which
specify the final curvature enhancement on the solution, these weights
vary between $0$ and $1$ with a value of $0$ makes no changes and
with values 1 applies a maximal change. The weights modify influence
zones where the Laplacian is applied as shown in the equation \ref{eq:Lineal_System_with_wp}.
Families of shapes that are generated may change substantially with
the weights of specific control points.

The models volume increases as the lambda is larger and negative,
this can be go against by a simple method of volume preservation.
In \cite{Desbrun1999} presents a simple method to resize the mesh
however the model suffers large displacements when $\lambda<-1.0$
or after multiple iterations. We propose the following solution: If
$v_{i}^{t+1}$ is a mesh vertex of $V^{t+1}$ in the $t+1$ iteration,
we define $\overline{v}$ as:

\begin{center}
$\overline{v}=\frac{1}{n}\underset{v_{i}\in V}{\sum}v_{i}$, 
\par\end{center}

$\overline{v}$ is the center of the mesh, $vol_{ini}$ is an initial
volume, and $vol_{t+1}$ is the volume at the iteration $t+1$, then
we define a scale factor.

\begin{center}
$\beta=\left(\frac{vol_{ini}}{vol_{t+1}}\right)^{\frac{1}{3}}$
\par\end{center}

The new vertices positions are:

\begin{center}
$v_{i\, new}^{t+1}=\beta\left(v_{i}^{t+1}-\overline{v}\right)+\overline{v}$
\par\end{center}

\subsection{Sculpting\label{sub:Sculpting}}

\begin{figure*}[t]
\includegraphics[width=1\textwidth]{images/cruz_lambda4}

\caption{\label{fig:Catmull_Clark}Left: Original Model, in green color model
with Catmull-Clark Subdivision. Models with Laplacian smoothing: $\lambda=60.0$,
$\lambda_{e}=12.0$ and $\lambda=90.0$, $\lambda_{e}=80.0$. Models
first filter with Laplacian smoothing $\lambda=60.0$, $\lambda_{e}=12.0$
and before applied curvature enhancing: $\lambda=-200.0$, $\lambda_{e}=12.0$
and $\lambda=-40.0$, $\lambda_{e}=-100.0$.}
\end{figure*}

We designed a new brush that allows the enhancement of the curvatures
of a polygon mesh in real time. Our brush works well with the stroke
method named \textsl{Drag Dot}  which allows to preview the changes
in the model before you release the mouse button or pen tablet, also
allows to move the mouse over the model to fit exactly where you want
to perform the enhancement of the curvature.

Brushes that perform similar work as inflate can create distortions
in the mesh and can also produce self-intersections in the mesh, as
these brushes only move the vertices along the normal and do not take
into account the global information. Whereas our method looks for
the best way to make inflation while preserving the global curvature
retaining the several shape and sharp features of the model.

Our method simplifies the work required for the enhancement, which
would be to use some different brushes for inflating and some other
to soften and styling. With our enhanced brush in one step can be
performed.

For the real-time brush it necessary that the Laplacian matrix is
constructed with the vertices that are within the radius  of the sphere
defined by the user, which reduces the size of the matrix to be processed,
the center of this sphere depends on the place where the user clicks
on the canvas and three-dimensional mesh place where the click is
projected.

Furthermore special handling is required for the vertices at the boundary
which have neighbors that are not within the radius of the brush.
These vertices must be marked as boundary and the curvature will not
be calculated for them, but must be present in the matrix so that
every vertex has their neighbors within the selection, this change
allows the results to be much smoother on the border. The Laplacian
matrix for sculpt mode as a matrix equation.

\begin{center}
$L\left(i,j\right)=\begin{cases}
-\frac{w_{ij}}{\underset{j\in N\left(v_{i}\right)}{\sum}w_{ij}} & \mbox{if }\left\Vert v_{i}-u\right\Vert <r\wedge\left\Vert v_{j}-u\right\Vert <r\\
0 & \mbox{if }\left\Vert v_{i}-u\right\Vert <r\wedge\left\Vert v_{j}-u\right\Vert \geq r\\
\delta_{ij} & \mbox{otherwise}
\end{cases}$
\par\end{center}

Where $v_{j}\in N\left(v_{i}\right)$, $u$ is the center of sphere
of radius $r$. The matrixes should remove rows and columns of vertices
index that are not within the radius.

\subsection{Subdivision surfaces\label{sub:Subdivision-surfaces}}

The Catmull-Clark subdivision transformation is used to smooth a surface
as the limit of a sequence of subdivision steps\cite{Stam1998}. This
process is governed by properties of B-splines curve from multivariate
spline theory\cite{Loop1987}. This method does a recursive subdivision
transformation that refines the model into a linear interpolation
that approximates a smooth surface. The smoothness of the model is
automatically guaranteed \cite{DeRose1998}.

Catmull-Clark subdivision surface methods generate smooth and continuous
models from a coarse model and produce results quickly due to the
simplicity of implementation, but with these methods is not easy to
make changes to the global curvature of the model. If we use Catmull-Clark
subdivision surfaces and curvature enhancement for modeling from coarse
models with few vertices can generate families of shapes changing
a single parameter, this would allow an artist to choose the model
from a similar set of options that best meets his/her needs without
having to change each one of the control vertices.

Our method allows the use of vertex weight paint over the control
points. The weights can be applied to a coarse model, then to perform
a Catmull-Clark subdivision where weights are interpolated, producing
weights with smooth changes in the zones of influence, so the obtained
curvature is much softer at those areas as shown in figure \ref{fig:Subdivision-Cups}.c.

In the equation \ref{eq:Lineal_System_with_wp}, $W_{p}$ is a diagonal
matrix with weights corresponding to each vertex. Weights at each
vertex produce a different solution for this reason the matrix is
placed in the diffusion equation, as families that are generated may
change substantially with weighted of specific control points.

\section{Results\label{sec:Results}}

In this section we describe the results of our curvature enhance method
with the extension of the Laplace Beltrami operator for hybrid quad/triangle
meshes with several example models (see figures \ref{fig:Spectrum},
\ref{fig:Subdivision-Cups}, \ref{fig:Catmull_Clark}, \ref{fig:TQLBO_test},
\ref{fig:camello_enhanced}, \ref{fig:Sculpt_Brush}, \ref{fig:Performance-sculpt},
\ref{fig:(a)Monkey}, \ref{fig:Animated_Camell}). We test the curvature
enhance with TQLBO method on a PC with AMD Quad-Core Processor @ 2.40
GHz and 8 GB RAM.

Figure \ref{fig:TQLBO_test} show the results when applying Laplace
Beltrami Operator TQLBO of equation (\ref{eq:TQLBO_Simple_Matrix})
in a model with a simple subdivision. In column (c) Laplacian smoothing
was applied to the model consists of only quads. In column (d) the
model was converted to triangles and then Laplacian smoothing was
applied. In column (e) the model is converted randomly some quads
into triangles and then Laplacian smoothing is applied, showing similar
results to those composed only of triangles or squares.

\begin{figure}
\includegraphics[width=1\columnwidth]{images/test_triangles_quads}

\caption{\label{fig:TQLBO_test}(a) Original Model. (b) Simple subdivision.
(c), (d) (e) Laplacian smoothing with $\lambda=7$ and 2 iterations:
(c) for triangles, (d) for quads, (e) for triangles and quads random
chosen.}
\end{figure}

Methods using Catmull-Clark subdivision surface and the enhancement
allows to modify the curvature that is obtained with the process of
subdivision as shown in figure \eqref{fig:Subdivision-Cups} this
test used a coarse model of a cup, after the model was performed subdivision
then was performed Laplacian smoothing and enhancement by modifying
the parameters $\lambda$ y $\lambda_{e}$ corresponding to the lambda
for rings and edges respectively. In figure \eqref{fig:Subdivision-Cups}.c,
\eqref{fig:Subdivision-Cups}.d shows also the use weight vertex groups
over coarse models with subdivision surfaces that allowed to generate
the weights for the new interpolated vertices, these new weights were
used for the enhancement obtained the 6 cups that are to the right
of the figure \eqref{fig:Subdivision-Cups}.d.

Laplacian smoothing applied with simple subdivision may produce similar
results to those obtained with Catmull-Clark whose models are on average
equal triangles as shown in figure \ref{fig:Catmull_Clark} green
model and that obtained with Laplacian smoothing $\lambda=60.0$,
$\lambda_{e}=12.0$, but also can modify the curvature obtained after
applying Catmull-Clark as shown in the three columns to the right
of the figure \ref{fig:Catmull_Clark}.

Figure \ref{fig:camello_enhanced} show the generation of different
versions of a camel according to parameter lambda. In the top row
you can see results of do curvature enhance over all model, as the
lambda becomes larger and negative in this model it inflates the convex
parts as shown in figure \ref{fig:Spectrum}. The larger $\lambda$
the larger enhancements on the model features. The bottom row of figure
\ref{fig:camello_enhanced} shows the use of weighted vertex groups,
which allows to specify which areas will be enhanced. On the left
the enhancement of the legs of the camel is shown which produces enhancement
of organic aspect, notice that the border is not distorted and smooth
in the union.

\begin{figure}
\includegraphics[width=1\columnwidth]{images/camello_enhanced2}

\caption{\label{fig:camello_enhanced}Top row: Original camel model in left.
Curvature enhancing with $\lambda=-30.0$, $\lambda=-100.0$, $\lambda=-400.0$.
Bottom row: Curvature enhancing with weight vertex group, $\lambda=-50.0$
and 2 iterations at legs, $\lambda=-200.0$ and 1 iteration in head
and neck.}
\end{figure}

Our method for enhancement of the silhouette features is predictable
and invariant under isometric transformations as those present in
some animations (see Figure \ref{fig:Animated_Camell}). This animation
shows some poses of a camel during a walk, enhancement is performed
in the neck and legs as shown in the bottom left part of the figure
\ref{fig:Animated_Camell}. Local modifications produced by pose interpolation
or animation rigging do not significantly affect the result as with
camel legs in each sample pose different flexion of the leg joints
enhancement camel keeps flesh-like shapes in the pattern produced
original by the artist. This is due to the diffusion process which
is subjected to the mesh so that small local changes are treated without
significantly affecting global solution. Our method is invariant of
rotation, this since depends exclusively on the normal field of the
mesh, wish is invariant under global rotations.

Figure \ref{fig:Sculpt_Brush} shows the use of curvature enhancement
brush for sculpting in real time. One pass was used with the brush
as shown in the figure with the blue and red radius. In figure \ref{fig:Sculpt_Brush}.b
camel shows how the leg self-intersecting and looks like two bubbles
stuck, the same happens to the fingers on the bottom of the figure.
Using silhouette features enhancement in figure \ref{fig:Sculpt_Brush}.c
we get better results sd the are retained shape of the silhouette
and the details of the fingers and leg. Similar results can be obtained
by an artist however it would take several steps and require the use
of several brushes. With curvature enhancement it only takes one step.
This new method can easily enhance organic features like muscles during
the sculpting process. In figure \ref{fig:Performance-sculpt} we
show the performance of curvature enhancement brush, in this experiment
we use three models with 12K, 40K and 164K vertices, this models were
sculpted with curvature enhance brush in each step the brush sphere
contain a variable number of vertices for processing. The processing
time for 800 vertices in the camel paw (40k model) only take 0.1 seconds,
for 2600 vertices in leg and neck (model 40k) take 0.5 seconds, these
times are good, because usually an artist sculpts a model for parts,
and each part is represented by an average of about 3000 vertices
in the models we use.

Tests with the Laplacian operator equation (\ref{eq:TQLBO_Simple_Matrix})
and its normalized version equation (\ref{eq:TQLBO-Normalized_Matrix}),
produce similar results if the triangles that compose the mesh are
the same size on average. The normalized version is more stable and
predictable because it is not divided by the area of the ring which
may be very small and cause problems during floating point calculations
as shown in figure \ref{fig:(a)Monkey}.c bottom row. The enhancement
of curvature of the model with the normalized Laplacian operator has
a more regular behavior. The model can be deformed in TQLBO normalized
version with large lambdas ($\lambda>400$) while it intersects itself
it produces no peaks. Figure \ref{fig:(a)Monkey} shows different
results due to the area of the triangles in the model. Triangles with
larger area have smaller enhancement (figure \ref{fig:(a)Monkey}.c
skull), and smaller triangles have larger enhancements (figure \ref{fig:(a)Monkey}.c
chin).

\begin{figure}
\includegraphics[width=1\columnwidth]{images/sculpt_brush}

\caption{\label{fig:Sculpt_Brush}Top row: (a) Leg Camel, (b) Inflate brush
for leg into blue circle, (c) Enhance curvature brush for leg into
red circle. Bottom row: (a) Hand, (b) Inflate brush for fingers into
blue circle, (c) Enhance curvature brush for fingers in red circle.}
\end{figure}

\begin{figure}
\includegraphics[width=1\columnwidth]{images/verts_per_second_sculpt}

\caption{\label{fig:Performance-sculpt}Performance of our dynamic curvature
enhance brush in terms of vertices sculpted per second. Three models
with 12K, 40K, 164K vertices used for sculpting in real time.}
\end{figure}

\begin{figure}
\includegraphics[width=1\columnwidth]{images/monkey}

\caption{\label{fig:(a)Monkey}(a) Top row: Original model scaled by 4. Bottom
row: Original Model (b) Top and bottom row: enhancing with Normalized-TQLBO
$\lambda=-50$ (c) Top and bottom row: enhancing with TQLBO $\lambda=-50$.}
\end{figure}

\begin{figure*}[t]
\includegraphics[width=1\textwidth]{images/camello_walk2}

\caption{\label{fig:Animated_Camell}Our method is pose insensitive. The enhanced
for the different poses are similar in terms of shape. Top row: Original
walk cycle camel model. Bottom row: Curvature enhancing with weight
vertex group, $\lambda=-400$ and $2$ iterations. }
\end{figure*}

\subsection{Implementation\label{sub:Implementation}}

Our method was implemented as a modifier for modeling and brush for
sculpting on the Blender software \cite{blender} with C and C++.
Working with Blender allowed us to test the method interactively against
others as Catmull-Clark, weight vertex groups and sculpting system
in Blender.

To improve the performance, we worked with Blender mesh struct visiting
each triangle or quad and adding its corresponding index in a list
that stores the sum of the Laplacian weights of the ring, in this
way we only had to visit two times the list of faces of the mesh list
and two times the list of the edges if the mesh was not closed. The
brush sculpting mode required to create a list that maps the index
of the selected vertices to a list from 1 to N where N is the number
of vertices selected and the number of rows in the linear system to
solve. This drastically reduced the calculations enabling real-time
processing. In the construction of our Laplacian matrix several index
were locked at vertices having faces areas or lengths edges with value
zero that can cause spikes and bad results.

Under this conditions matrix at the equation\ref{eq:TQLBO_Simple_Matrix}
is sparse since the number of neighbors per vertex corresponding to
the number of data per row is small compared to the total number of
vertices in the mesh. To solve the linear system equation \ref{eq:Lineal_System_with_wp}
was used OpenNL which is a a library for solving sparse linear system.

\section{Conclusion and future work\label{sec:Conclusion-and-future}}

This work presented an extension of the Laplace Beltrami operator
for hybrid quad/triangle meshes that can be used in production environments
without modification that provides results similar to those obtained
by working only with triangles or quads.

We introduced a new way to perform silhouettes enhancement in a mesh
for modeling or sculpting in a few steps which allows the modification
of the curvature of a model while preserving its overall shape. We
introduce a new method of modeling and show some of its possible uses.
This method behaves in predictable ways which facilitate the learning
process, and works well with isometric transformations opening the
possibility of introducing it on the process of animation.

We show that this tool could work in the early stages in where coarse
models are used, allowing to modify the curvature generated by the
Catmull-Clark subdivision surfaces, avoiding the edition of the vertices
with the change of a few parameters.

As future work we analyze theoretically the relationship between the
Catmull-Clark subdivision surfaces and Laplacian smoothing because
in some cases can produce very similar results. But the subdivision
surface is a fast method thereby reducing computation times for calculate
curvature in the mesh.

\section*{Acknowledgments}

We wold like to thank  anonymous friends for their support of our
research.

This work was supported in part by the Blender Foundation, Google
Summer of code program at 2012.

Livingstone elephant model is provided courtesy of INRIA and ISTI
by the AIM@SHAPE Shape Repository. Hand model is courtesy of the FarField
Technology Ltd. Camel model by Valera Ivanov is licensed under a Creative
Commons Attribution 3.0 Unported License. Dinosaur and Monkey models
are under public domain, courtesy of Blender Foundation.

\bibliographystyle{acmsiggraph/acmsiggraph}
\bibliography{template}

\end{document}