### 2015-04-12 trunk/documents/thesis/thesis_LO_QT.tex

%% LyX 2.1.0 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,letterpaper,english,12pt,english,openany,letterpaper,pagesize]{scrbook}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\setlength{\parskip}{\medskipamount}
\setlength{\parindent}{0pt}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{esint}
\onehalfspacing

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\pdfpageheight\paperheight
\pdfpagewidth\paperwidth

\newcommand{\lyxmathsym}[1]{\ifmmode\begingroup\def\b@ld{bold}
  \text{\ifx\math@version\b@ld\bfseries\fi#1}\endgroup\else#1\fi}

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%\documentclass[12pt,spanish,fleqn,openany,letterpaper,pagesize]{scrbook}

%\usepackage[ansinew]{inputenc}
%\usepackage[spanish]{babel}
\usepackage{fancyhdr}
\usepackage{epsfig}
\usepackage{epic}
\usepackage{eepic}
\usepackage{amsmath}
\usepackage{threeparttable}
\usepackage{amscd}
\usepackage{here}
\usepackage{graphicx}
\usepackage{lscape}
\usepackage{tabularx}
\usepackage{subfigure}
\usepackage{longtable}

\usepackage{rotating} %Para rotar texto, objetos y tablas seite. No se ve en DVI solo en PS. Seite 328 Hundebuch
                        %se usa junto con \rotate, \sidewidestable ....

\renewcommand{\theequation}{\thechapter-\arabic{equation}}
\renewcommand{\thefigure}{\textbf{\thechapter-\arabic{figure}}}
\renewcommand{\thetable}{\textbf{\thechapter-\arabic{table}}}

\pagestyle{fancyplain}%\addtolength{\headwidth}{\marginparwidth}
\textheight22.5cm \topmargin0cm \textwidth16.5cm
\oddsidemargin0.5cm \evensidemargin-0.5cm%
\renewcommand{\chaptermark}[1]{\markboth{\thechapter\; #1}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\; #1}}
\lhead[\fancyplain{}{\thepage}]{\fancyplain{}{\rightmark}}
\rhead[\fancyplain{}{\leftmark}]{\fancyplain{}{\thepage}}
\fancyfoot{}
\thispagestyle{fancy}%

\addtolength{\headwidth}{0cm}
\unitlength1mm %Define la unidad LE para Figuras
%\mathindent0cm %Define la distancia de las formulas al texto,  fleqn las descentra
\marginparwidth0cm
\parindent0cm %Define la distancia de la primera linea de un parrafo a la margen

%Para tablas,  redefine el backschlash en tablas donde se define la posici\'{o}n del texto en las
%casillas (con \centering \raggedright o \raggedleft)
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

%Espacio entre lineas
\renewcommand{\baselinestretch}{1.1}

%Neuer Befehl f\"{u}r die Tabelle Eigenschaften der Aktivkohlen
\newcommand{\arr}[1]{\raisebox{1.5ex}[0cm][0cm]{#1}}

%Neue Kommandos
\usepackage{resources/Befehle}

%Trennungsliste
\hyphenation {Reaktor-ab-me-ssun-gen Gas-zu-sa-mmen-set-zung
Raum-gesch-win-dig-keit Durch-fluss Stick-stoff-gemisch
Ad-sorp-tions-tem-pe-ra-tur Klein-schmidt
Kohlen-stoff-Mole-kular-siebe Py-rolysat-aus-beu-te
Trans-port-vor-gan-ge}

%\pagenumbering{roman}
%\let\myTOC\tableofcontents
%\renewcommand\tableofcontents{%
%\myTOC
%\clearpage
%\pagenumbering{arabic}
%}

\makeatother

\usepackage{babel}
\begin{document}
\pagenumbering{roman}
\begin{center}
\begin{figure}
\centering%
%\includegraphics[width=0.4\columnwidth]{resources/logo_un.png}
%\includegraphics[width=0.4]{logo_un.png}
\end{figure}
\thispagestyle{empty}

\vspace{2cm}

\textbf{\huge{}An Adapted Laplacian Operator For Hybrid Quad/Triangle
Meshes}{\huge \par}

\vspace{6cm}

\textbf{\large{}Alexander Pinzón Fernández}{\large \par}

\vspace{6cm}

Universidad Nacional de Colombia

Facultad de Ingeniería, Departamento de Sistemas e Industrial

{\small{}Grupo de Investigación CIM@LAB}{\small \par}

Bogotá, Colombia

2014

\end{center}

\newpage{}

\begin{center}
\thispagestyle{empty} 
\vspace*{0cm}

\textbf{\huge{}An Adapted Laplacian Operator For Hybrid Quad/Triangle
Meshes}{\huge \par}

\vspace{2.5cm}

\textbf{\large{}Alexander Pinzón Fernández}{\large \par}

\vspace{2.5cm}

{\small{}A thesis submitted in partial fulfillment of the requirements
for the degree of:}{\small \par}

\textbf{Master in Systems Engineering and Computer Science}

\vspace{2cm}

{\small{}Advisor:}\\
{\small{}Eduardo Romero Castro , Ph.D.}{\small \par}

{\small{}\vspace{1.5cm}
}{\small \par}

{\small{}Research Area:}\\
{\small{}Computer Graphics}{\small \par}

{\small{}\vspace{2cm}
}{\small \par}

{\small{}Universidad Nacional de Colombia}\\
{\small{}Facultad de Ingeniería, Departamento de Sistemas e Industrial}\\
{\small{}Grupo de Investigación CIM@LAB}\\
{\small{}Bogotá, Colombia}\\
{\small{}2014}{\small \par}

\end{center}
\newpage 
\thispagestyle{empty}

\textbf{Dedicación}

{\small{}\vspace{4cm}
}{\small \par}

\noindent \begin{flushright}
\begin{minipage}[t]{8cm}%
{\small{}A Beatriz y Campo Elias mis padres que siempre me dieron
la libertad de escoger, con su comprensión y apoyo me permitieron
dedicar mi tiempo a la ciencia.}{\small \par}

{\small{}\vspace{1cm}
}{\small \par}

{\small{}A mis padres}{\small \par}

{\small{}\vspace{1cm}
}{\small \par}

Beatriz Fernandez Vargas

Campo Elias Pinzón Rojas%
\end{minipage}
\par\end{flushright}

\newpage
\thispagestyle{empty}

\textbf{\LARGE{}Acknowledgment}{\LARGE \par}

\addcontentsline{toc}{chapter}{\protect\numberline{}Acknowledgement}

\vspace{2cm}

I wold like to thank my advisor professor Eduardo Romero and the CIM\&LAB
research group for their support in this thesis.

{\small{}\vspace{1cm}
}{\small \par}

I would like to thank all the workers in Colombia who fund the public
education with their work.

{\small{}\vspace{1cm}
}{\small \par}

This work was supported in part by the Blender Foundation, Google
Summer of code program 2012 and 2013.

\newpage{}

\textbf{\LARGE{}Abstract}{\LARGE \par}

\addcontentsline{toc}{chapter}{\protect\numberline{}Abstract}

In the last two decades three-dimensional modeling methods used by
artists have been evolving and developing rapidly thanks to the use
of vector operators of differential geometry such as the Laplacian
operator. This operator allows to model in a simple way the behavior
of complex applications such as noise reduction, enhancement, remeshing,
UV mapping, posing and skeletonization, among others. The Laplacian
operator is theoretically defined in a continuous and smooth domain
named manifold. In practice manifolds are often approximated by discrete
polygon meshes composed by triangles and quadrangles which represent
the real world three-dimensional objects with which the artists work.
In these meshes spectral structure is calculated using a discrete
Laplacian operator, i.e. the discrete version of the Laplacian operator
given by Pinkall in 1993. This approach only worked with triangle
meshes. In 2011 Xiong extended the operator to work exclusively with
quad meshes. This thesis proposes an original extension of the Laplacian
operator that allows working with hybrid meshes composed by triangles
and quadrangles.

Along with the operator, this work presents new sculpting and modeling
applications based on enhancement. Additionally applications on subdivision
surfaces using smoothing, mesh posing using differential coordinates
and skeletonization using iterative contraction are develop. This
series of applications demonstrates the quality, predictability and
flexibility of the proposed operator.

The proposed operator was successfully used in new software tools
in real production environment within 3D computer graphics software
Blender. Currently these tools are available as open source software.

\textbf{\LARGE{}Resumen}{\LARGE \par}

En las dos últimas décadas los métodos de modelado tridimensional
utilizadas por los artistas han ido evolucionando y desarrollándose
rápidamente, en parte gracias al uso de operadores vectoriales de
geometría diferencial, como el operador de Laplace. Este operador
permite modelar de una manera sencilla el comportamiento de aplicaciones
complejas tales como la reducción de ruido, realce, remallado, mapeado
UV, posado y esqueletonización, entre otros. Este operador Laplaciano
es teóricamente definido en un dominio continuo y suave llamado variedad,
las variedades son a menudo aproximadas por mallas discretas de polígonos
compuestas por triángulos y cuadrángulos que a su vez representan
objetos tridimensionales del mundo real que los artistas trabajan.
En estas mallas se calcula la estructura espectral con el uso de algún
operador Laplaciano discreto, la versión discreta del operador Laplaciano
propuesta por Pinkall en el 1993 trabaja únicamente con mallas compuestas
por triángulos, y la de Xiong en el 2011 trabaja exclusivamente con
cuadrángulos. Esta tesis propone una extensión original del Operador
Laplaciano que permite trabajar con mallas híbridas compuestas por
triángulos y cuadrángulos.

Junto con el operador, este trabajo presenta nuevas aplicaciones en
esculpido y modelamiento con base en el realce, aplicaciones en subdivisión
de superficies con el uso de suavizado, posado de mallas con el uso
de coordenadas diferenciales y esqueletonización usando contracción
iterativa. Esta serie de aplicaciones demuestra la calidad, predictibilidad
y flexibilidad del operador propuesto.

El operador propuesto fue usado con exitoso en las nuevas herramientas
del software para gráficos 3D por computadora Blender. Actualmente
estas herramientas están disponibles como programas de código abierto.

{\small{}\vspace{1cm}
}{\small \par}

\textbf{\small{}Keywords: laplacian operator; smooth; enhance; sculpting;
subdivision surface}{\small \par}

\tableofcontents{}

\listoffigures

\listoftables

\pagenumbering{arabic}

\chapter{Introduction}

The discrete versions of the Laplace Beltrami Operator have been used
in the last years for the development of new geometric modeling tools.
Pinkall \cite{Pinkall1993} introduced the cotangent version of the
Laplace operator, that allowed finding the minimal surface when computing
a discrete harmonic map with the Laplacian operator. This version
has been widely studied and applied in various problems of computer
geometric modeling. This type of operator was defined over manifolds,
i.e. continuous domains homeomorphic to $\mathbb{R}^{n}$ that are
represented in pratice by polygon meshes. These polygons are generally
composed of triangles and quadrangles. While working with Laplacian
operator in this hybrid composition is not a mathematical challenge,
most research only uses meshes composed by triangles \cite{Pinkall1993,Desbrun1999,Meyer2003,Sorkine2004,Au2008,Belkin2008,Hahne2006}.
In recent studies \cite{Liu2008,Xiong2011}, the Laplacian operator
may be use in meshes composed exclusively by quadrangles. However,
from artistic point of view, the topology and the way the edges, triangles
and quadrangles are distributed directly affects the processes of
animation, interpolation, texturing, etc., as discussed by \cite{Mullen2007}
who uses a manual connection of a pair of vertices to perform animation
processes and interpolation. It is then of paramount importance to
develop operators that easily interact with such meshes, eliminating
the need of preprocessing the mesh to convert it to triangles and
change the original topology.

Presently there are available modeling techniques able to generate
a variety of realistic shapes \cite{Botsch2006}. Editing techniques
have evolved from affine transformations to advanced tools like sculpting
\cite{Coquillart1990,Galyean1991,Stanculescu2011}, editing, creation
from sketches \cite{Igarashi1999,Gonen2012}, and complex interpolation
techniques \cite{Sorkine2004,Zhou2005}. Catmull-Clark based methods
however require to interact with a minimum number of control points
for any operation to be efficient, or in other words, a unicity condition
is introduced by demanding a smooth surface after any of these shape
operations. Hence, traditional modeling methods for subdividing surfaces
from coarse geometry have become widely popular \cite{Catmull1978,Stam1998}.
These works have generalized a uniform B-cubic spline knot insertion
to meshes. Some of these adds some type of control; for instance with
the use of creases to produce sharp edges \cite{DeRose1998} or the
modification of some vertex weights to locally control the zone of
influence \cite{Biermann2000}. Nevertheless  these methods are difficult
to use as they require a large number of parameters and a very tedious
customization.

On the other hand, the proposed applications requires a single parameter
that controls the global curvature, which is used to maintain realistic
shapes, creating a family of different versions of the same object
and therefore preserving the detail of the original model and a realistic
appearance.

The shape inflation and shape exaggeration can thus be used as a type
of brush in the sculpting process. When inflating a shape with other
brushes the former ends up losing detail when moving vertices \cite{Stanculescu2011}.
In contrast, the presented enhance method inflates a mesh by moving
the vertices towards the reverse curvature direction, conserving the
shape and sharp features of the model.

\paragraph{Contributions}

This work presents an extension of the Laplace Beltrami Operator for
hybrid quad/triangle meshes that have a larger mesh functionality
spectrum than common triangular or quadrangular meshes. The method
eliminates the need of preprocessing and allows preservation of the
original topology. Along with this operator, we propose a method to
generate a family of parameterized shapes, in a robust and predictable
way. This method enables customization of the smoothness and curvature
obtained during the subdivision surfaces process. Finally, a new brush
for inflating the silhouette mesh features in modeling and sculpting
is proposed.

The work is organized as follows: Chapter \ref{chap:Mathematical-foundation}
presents works related to the Laplacian operator, applications in
digital sculpting, deformation, and offsetting methods for polygonal
meshes. It also describes the theoretical framework of the Laplacian
operator for polygon meshes; In chapter \ref{chap:SibgrapiPaper},
we show the extension of the Laplacian Operator for hybrid meshes
and the applications of shape inflation ,subdivision of surfaces and
sculpting; In chapter \ref{sub:Laplacian-Smooth}, we present an application
for mesh smoothing of the the Laplacian operator extension proposed
implemented in a well known software for computer modelling. Finally
in chapter \ref{sub:Laplacian-Deform}, we present a successfull application
for mesh deformation and model re-posing based on differential coordinates
and the adaptation of this method to work with our Laplacian operator
extension.

\chapter{Mathematical Foundation and Background\label{chap:Mathematical-foundation}}

This chapter studies basic mathematical foundation on differential
geometry to understand the differential operators and the Laplace
Beltrami operator.

The differential geometry studies curvatures and geodesics \cite{Hahne2006}.
These differential operators show a deep relationship between the
geometry (curvatures, geodesics) and topology of the manifold. They
have been commonly used in computer geometric modelling applications
over the last years \cite{Rosenberg1997,Alexa2011}.

\section{Related work}

Many tools have been developed for modeling based on the Laplacian
mesh processing. These tools preserve the surface geometric details
when using laplacian operators for different processes such as smoothing,
enhancing, free-form deformation, fusion, morphing and other applications
\cite{Ranjan2012}.

The most used discretization of Laplace Beltrami operator $\Delta_{\mbox{\ensuremath{\Omega}}}$
over a triangulated mesh $\Omega$ was proposed by Pinkall \cite{Pinkall1993}.

\noindent \begin{center}
$\Delta_{\mbox{\ensuremath{\Omega}}}\left(u\right)=\frac{1}{2}\sum_{j\in N_{1}\left(i\right)}\left(\cot\alpha_{j}+\cot\beta_{j}\right)\left(x_{i}-x_{j}\right)$
\par\end{center}

Where $N_{1}$is the 1-ring neighborhood, $\alpha$ and $\beta$ are
the opposite angles to edge between vertex $i$ and vertex $j$. In
this work the discrete laplacian operator is used to find the minimal
surface based on minimization energy strategy using the Dirichlet's
energy of the function $u$ over a manifold represented by triangulated
mesh $\Omega$.

\noindent \begin{center}
$E_{D}\left(u\right)=\frac{1}{2}\int_{\Omega}\bigl|\nabla u\bigr|^{2}$
\par\end{center}

Taubin \cite{Taubin1995} treats for the first time the problem of
noise reduction in digital polygonal meshes from signal processing
point of view. He extends Fourier analysis to signals defined on polygonal
meshes and observes that Fourier transform is a decomposition of the
signal into eigenvectors of the Laplacian operator and reconstructs
the signal with a linear combination of these eigenvectors. Desbrun
et al. \cite{Desbrun1999} consider the same approach of Taubin, but
they use a curvature normal ($\bar{\kappa}\mathbf{n}$) based on cotangent
Laplace operator version for noise reduction over a diffusion process.
This is the most famous and popular Laplace Beltrami operator discretization
\cite{Liu2008}, many works for mesh smoothing and fairing have been
developed based on his Laplace Beltrami Operator (LBO) discretization
\cite{Desbrun2000,Meyer2003,Schneider2001,Nealen2006}:

\noindent \begin{center}
$\frac{\partial x_{i}}{\partial t}=-\bar{\kappa_{i}}\mathbf{n_{i}}$
\par\end{center}

\noindent \begin{center}
$-\bar{\kappa_{i}}\mathbf{n_{i}}=\frac{1}{A}\sum_{j\in N_{1}\left(i\right)}\left(\cot\alpha_{j}+\cot\beta_{j}\right)$
\par\end{center}

where $A$ is the area surrounding vertex $i$. This laplacian operator
$L$ was used to reduce the noise in a mesh $X$ over a diffusion
process.

The convergence of the Laplace Beltrami operator has been very important
in fields such as numerical analysis, given its implications in the
simulation process and geometric partial differentials equations.
Xu et al. \cite{Xu2004} established the convergence of several discrete
Laplace Beltrami operators over triangulated meshes with numerical
results that support the theoretical analysis. Over quadrilateral
meshes Liu et al. \cite{Liu2008} present a discrete Laplace Beltrami
Operator based on a bilinear interpolation and its convergence over
meshes composed only by quads.

In the work of Sorkine \cite{Sorkine2006} the laplacian operator
was used to re-pose a mesh while preserving geometry details of the
surface. The details were stored in differential coordinates $\delta_{i}$
for every vertex $v_{i}$.

\noindent \begin{center}
$\delta_{i}=\sum_{j\in N_{1}\left(i\right)}\frac{1}{2}\left(\cot\alpha_{j}+\cot\beta_{j}\right)\left(v_{i}-v_{j}\right)$
\par\end{center}

The differential coordinates represent the difference between the
absolute coordinate of $v_{i}$and the center of mass of its immediate
neighbors.

Offset methods for polygon meshing, based on the curvature defined
by the Laplace Beltrami operator, have been developed. These methods
adjust the shape offset by a constant distance, with high precision.
Nevertheless, these methods fail to conserve sufficient detail because
of the smoothing, a crucial issue which depends on the offset size
\cite{Zhuo2012}. In volumetric approaches, in case of point-based
representations, the offset boundary computation is based on the distance
field and therefore when calculating such offset, the topology of
the model may be different to the original \cite{Chen2011}.

Gal et al. \cite{Gal2009} propose automatic feature detection and
shape edition with feature inter-relationship preservation. They define
salient surface features like ridges and valleys, characterized by
their first and second order curvature derivatives (see \cite{Ohtake2004})
and angle-based thresholds. Likewise, curves have been also classified
as planar or non-planar, approximated by lines, circles, ellipses
and other complex shapes. In each case, the user defines an initial
change over several features which is propagated towards other features,
based on the classified shapes and the inter-relationships between
them. This method works well with objects that have sharp edges, composed
of basic geometric shapes such as lines, circles or ellipses. However,
the method is very limited when models are smooth since it cannot
find the proper features to edit.

Digital sculpting has been traditionally approached either under a
polygonal representation or a voxel grid-based method. Brushes for
inflation operations only depend on the vertex normal \cite{Stanculescu2011}.
In grid-based sculpting, some other operations have allowed to add
or remove voxels since production of polygonal meshes require a processing
of isosurfaces from a volume \cite{Galyean1991}. The drawback comes
from the difficulty of maintaining the surface details during larger
scale deformations.

In the literature, several studies have described the skeleton extraction
systems and different metrics that identify appropriate methods given
a specific application \cite{Attali2007}. One of the best methods
reported in the literature for the extraction of the skeleton is the
Laplacian smoothing method given its advantages of homotopy representation
and hierarchical connections between parts. The skeleton extraction
methods permit simplifying the dimension of the object preserving
the topological structure \cite{Cornea2007}, Au et. al. \cite{Au2008}
present a skeleton extraction method based on iterative smoothing-contraction.
In this method several constraints are used to warrant that the process
converges to a skeleton formed by branches and joints. The constraints
are based on laplacian operator; the low frequencies of the mesh are
preserved with the use of an attractor to the original mesh, while
the iterative smoothing process removes high frequencies.

\section{Manifolds}

A manifold is a topological space $M$ with the following properties:

If $x\in M$, then there is some neighborhood $N\left(x\right)$ and
some integer $n\geq0$ such that $N(x)$ is homeomorphic to $\mathbb{R}^{n}$
\cite{Spivak1-1999}.

Our work is related to manifolds that represent a surface in an three-dimensional
Euclidean Space. This manifolds are homeomorphic to $\mathbb{R}^{2}$.

The manifolds are represented by polygonal meshes with points connected
by triangles and quads.

\section{Laplace Operator}

In computer graphics a manifold is often approximated by a discrete
mesh \cite{Ranjan2012}, then is necessary to define a discrete laplacian
operator that act on functions defined on such meshes.

Consider a smooth compact manifold $M$ of dimension $m$ isometrically
embedded in a Euclidean space $\mathbb{R}^{d}$.

Given a twice continuously differentiable function $f\in C^{2}\left(M\right)$,
let $\nabla_{\mbox{M}}f$ denote the gradient vector field of $f$
on $M$.

The Laplace-Beltrami operator $\Delta_{\mbox{M}}$ of $f$ is defined
as the divergence of the gradient; that is \cite{Spivak1-1999},

\noindent \begin{center}
$\frac{\partial^{2}f}{\partial x^{2}}+\frac{\partial^{2}f}{\partial y^{2}}+\frac{\partial^{2}f}{\partial z^{2}}=0$
\par\end{center}

\noindent \begin{center}
$\nabla_{\mbox{M}}^{2}f=\Delta_{\mbox{M}}f=0$
\par\end{center}

\noindent 
\begin{equation}
\Delta_{\mbox{M}}f=div\left(\nabla_{\mbox{M}}f\right)
\end{equation}

\subsection{Discrete Laplace Operator Setting}

Discrete Laplacian operators are linear operators that act on functions
defines on meshes. These functions are specified by their values at
the vertices.

Thus, if a mesh $M$ has $n$ vertices, then functions on $M$ will
be represented by vectors with $n$ components and a mesh Laplacian
will be described by an $n\times n$ matrix \cite{Ranjan2012}.

The Laplacian operator locally takes the difference between the value
of a function at a vertex and a weighted average of its values at
the first-order or one-ring neighbor vertices, then a Laplacian matrix
$L$ has a local form given by

\noindent \begin{center}
$L\left(f\right)_{i}=b_{i}^{-1}\underset{j\in N\left(i\right)}{\sum}w_{ij}\left(f_{i}-f_{j}\right)$
\par\end{center}

Where $w_{ij}$are the weights between the vertex $i$ and the vertex
$j$. $b_{i}^{-1}$ are the factors depending of the boundary region
over vertex $i$. $N\left(i\right)$ are the neighbors that share
a edge with vertex $i$.

\chapter{Shape Inflation With an Adapted Laplacian Operator For Hybrid Quad/Triangle
Meshes\label{chap:SibgrapiPaper}}

\noindent \begin{center}
Alexander Pinzón, Eduardo Romero\\
Cimalab Research Group\\
Universidad Nacional de Colombia\\
Bogota-Colombia\\
Email: apinzonf@unal.edu.co, edromero@unal.edu.co
\par\end{center}

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/spectrum}

\protect\caption{\label{PAPER:fig:Spectrum}A set of 48 successive shapes enhanced,
from $\lambda=0.0$ in blue to $\lambda=-240.0$ in red, with steps
of $-5.0$.}
\end{figure}

This paper was accepted and presented in the Brazilian Symposium on
Computer Graphics and Image Processing SIBGRAPI 2013. The conference
was held on August 5-8 of 2013, in the city of Arequipa, Perú \cite{Pinzon2013}.

\section*{Abstract}

This paper proposes a novel modeling method for a hybrid quad/triangle
mesh that allows to set a family of possible shapes by controlling
a single parameter, the global curvature. The method uses an original
extension of the Laplace Beltrami operator that efficiently estimates
a curvature parameter which is used to define an inflated shape after
a particular operation performed in certain mesh points. Along with
the method, this work presents new applications in sculpting and modeling,
with subdivision of surfaces and weight vertex groups. A series of
graphics examples demonstrates the quality, predictability and flexibility
of the method in a real production environment with software Blender.
\begin{description}
\item [{keywords-}] laplacian smooth; curvature; sculpting; subdivision
surface
\end{description}

\section{Introduction}

Over the last years several, modeling techniques able to generate
a variety of realistic shapes, have been developed \cite{Botsch2006}.
Editing techniques have evolved from affine transformations to advanced
tools like sculpting \cite{Coquillart1990,Galyean1991,Stanculescu2011},
editing, creation from sketches \cite{Igarashi1999,Gonen2012}, and
complex interpolation techniques \cite{Sorkine2004,Zhou2005}. Catmull-Clark
based methods however require to interact with a small number of control
points for any operation to be efficient, or in other words, a unity
 condition is introduced by demanding a smooth surface after any of
these shape operations. Hence, traditional modeling methods for subdividing
surfaces from coarse geometry have become widely popular \cite{Catmull1978,Stam1998}.
These works have generalized a uniform B-cubic spline knot insertion
to meshes, some of them adding some type of control, for instance
with the use of creases to produce sharp edges \cite{DeRose1998},
or the modification of some vertex weights to locally control the
zone of influence \cite{Biermann2000}. Nevertheless these methods
are difficult to deal with since they require a large number of parameters
and a very tedious customization. Instead, the presented method requires
a single parameter that controls the global curvature, which is used
to maintain realistic shapes, creating a family of different versions
of the same object and therefore preserving the detail of the original
model and a realistic appearance.

Interest in meshes composed of triangles and quads has lately increased
because of the flexibility of modeling tools such as Blender 3D \cite{blender}.
Nowadays, many artists use a manual connection of a couple of vertices
to perform animation processes and interpolation \cite{Mullen2007}.
It is then of paramount importance to develop operators that easily
interact with such meshes, eliminating the need of preprocessing the
mesh to convert it to triangles. The shape inflation and shape exaggeration
can thus be used as such brush in the sculpting process, when inflating
a shape since current brushes end up by losing detail when moving
vertices \cite{Stanculescu2011}. In contrast, the presented method
inflates a mesh by moving the vertices towards the reverse curvature
direction, conserving the shape and sharp features of the model.

\subsection*{Contributions }

This work presents an extension of the Laplace Beltrami operator for
hybrid quad/triangle meshes, representing a larger mesh spectrum from
what has been presented so far. The method eliminates the need of
preprocessing and allows preservation of the original topology. Likewise,
along with this operator, it is proposed a method to generate a family
of parametrized shapes, in a robust and predictable way. This method
enables customization of the smoothness and curvature, obtained during
the subdivision surfaces process. Finally, it is proposed a new brush
for inflating the silhouette mesh features in modeling and sculpting.

This work is organized as follows: Section \ref{sec:Related-work}
presents works related to the Laplacian mesh processing, digital sculpting,
and offsetting methods for polygonal meshes; In section \ref{sec:Laplacian-Smooth},
it is described the theoretical framework of the Laplacian operator
for polygon meshes; In section \ref{PAPER:sec:Proposed-Method}, it
is presented the method for shape inflation and applications of subdivision
of surfaces and sculpting; finally some Laplacian operator results,
to hybrid quad/triangle meshes are graphically shown as well as results
of the shape inflation applications in sculpting, subdivision and
modeling.

\section{Related work\label{sec:Related-work}}

Many tools have been developed for modeling, based on the Laplacian
mesh processing. Thanks to the advantages of the Laplacian operator,
these different tools preserve the surface geometric details when
using them for different processes such as free-form deformation,
fusion, morphing and other applications \cite{Sorkine2004}.

Offset methods for polygon meshing, based on the curvature defined
by the Laplace Beltrami operator, have been developed. These methods
adjust the shape offset by a constant distance, with enough precision.
Nevertheless, these methods fail to conserve sufficient detail because
of the smoothing, a crucial issue which depends on the offset size
\cite{Zhuo2012}. In volumetric approaches, in case of point-based
representations, the offset boundary computation is based on the distance
field and therefore when calculating such offset, the topology of
the model may be different to the original \cite{Chen2011}.

\cite{Gal2009} proposes automatic feature detection and shape edition
with feature inter-relationship preservation. They define salient
surface features like ridges and valleys, characterized by their first
and second order curvature derivatives, see \cite{Ohtake2004}, and
angle-based threshold. Likewise, curves have been also classified
as planar or non-planar, approximated by lines, circles, ellipses
and other complex shapes. In such case, the user defines an initial
change over several features which are propagated towards other features,
based on the classified shapes and the inter-relationships between
them. This method works well with objects that have sharp edges, composed
of basic geometric shapes such as lines, circles or ellipses. However,
the method is very limited when models are smooth since it cannot
find the proper features to edit.

Digital sculpting has been traditionally approached either under a
polygonal representation or a voxel grid-based method. Brushes for
inflation operations only depend on the vertex normal \cite{Stanculescu2011}.
In grid-based sculpting, some other operations have allowed to add
or remove voxels since production of polygonal meshes require a processing
of isosurfaces from a volume \cite{Galyean1991}. The drawback comes
from the difficulty of maintaining the surface details during larger
scale deformations.

\section{Laplacian Smooth\label{sec:Laplacian-Smooth}}

Computer objects, reconstructed from the real world, are usually noisy.
Laplacian Smooth techniques allow a proper noise reduction on the
mesh surface with minimal shape changes, while still preserving a
desirable geometry as well as the original shape.

Many smoothing Laplacian functionals regularize the surface energy
by controlling the total surface curvature $S$.

\noindent \begin{center}
$E\left(S\right)=\int_{S}\kappa_{1}^{2}+\kappa_{2}^{2}dS$
\par\end{center}

Where $\kappa_{1}$ and $\kappa_{2}$ are the two principal curvatures
of the surface $S$.

\subsection{Gradient of Voronoi Area}

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/voronoi_region}

\protect\caption{\label{fig:voronoi_region-1}Area of the Voronoi region around $v_{i}$
in dark blue.$v_{j}$ belong to the first neighborhood around $v_{i}$.
$\alpha_{j}$ and $\beta_{j}$ opposite angles to edge $\protect\overrightarrow{v_{j}-v_{i}}$. }
\end{figure}

Consider a surface $S$ composed of a set of triangles around vertex
$v_{i}$. Let us define the \textsl{Voronoi region} of $v_{i}$ as
show in figure \ref{fig:voronoi_region-1}, The area change produced
by the movement of $v_{i}$ is called the gradient of \textsl{Voronoi
region} \cite{Pinkall1993,Desbrun1999,Meyer2003}.

\begin{equation}
\nabla A=\frac{1}{2}\underset{j}{\sum}\left(\cot\alpha_{j}+\cot\beta_{j}\right)\label{PAPER:eq:eq_gradient_voronoi_area}
\end{equation}

If the gradient in equation (\ref{PAPER:eq:eq_gradient_voronoi_area})
is normalized by the total area of the $1$-ring neighborhood around
$v_{i}$, the \textsl{discrete mean curvature normal} of a surface
$S$ is obtained, as shown in equation (\ref{PAPER:eq:discrete_mean_curvature_normal}).

\begin{equation}
2\kappa\mathbf{n}=\frac{\nabla A}{A}\label{PAPER:eq:discrete_mean_curvature_normal}
\end{equation}

\subsection{Laplace Beltrami Operator}

The \textsl{Laplace Beltrami operator} LBO noted as $\triangle$ is
used for measuring the mean curvature normal to the Surface $S$ \cite{Pinkall1993}.

\begin{equation}
\triangle S=2\kappa\mathbf{n}\label{PAPER:eq:def_LBO}
\end{equation}

The LBO has desirable properties: the LBO points out to the direction
of the minimal surface area.

\section{Proposed Method\label{PAPER:sec:Proposed-Method}}

This method exaggerates a shape using a Laplacian smoothing operator
in the reverse direction, i.e., the new shape is a modified version
in which those areas with larger curvature are magnified. The operator
amounts to a generator of a set of models which conserves the basic
silhouette of the original shape. In addition, the presented approach
can be easily mixed with traditional or uniform subdivision of surfaces.
This method is based on an original extension of the Laplace Beltrami
operator for hybrid quad/triangle meshes, mixing arbitrary types of
meshes, exploiting the basic geometrical relationships and ensuring
good results with few algorithm iterations.

\subsection{Laplace Beltrami Operator for Hybrid Quad/Triangle Meshes TQLBO\label{PAPER:sub:Laplace-Beltrami-Operator}}

Given a mesh $M=\left(V,Q,T\right)$, with vertices $V$, quads $Q$,
triangles $T$. The area of $1$-ring neighborhood $A\left(v_{i}\right)$
corresponds to a sum of the quad faces $A\left(Q_{v_{i}}\right)$
and the areas of the triangular faces $\ensuremath{A\left(T_{v_{i}}\right)}$
adjacent to vertex $\ensuremath{v_{i}}$.

\noindent \begin{center}
$\ensuremath{A\left(v_{i}\right)=A\left(Q_{v_{i}}\right)+A\left(T_{v_{i}}\right)}$
\par\end{center}

\begin{figure}[H]
\includegraphics[width=1\columnwidth]{resources/figs/quad_xiong}

\protect\caption{\label{PAPER:fig:quad_xiong}$t_{j1}^{*}\equiv\vartriangle v_{i}v_{j}v_{j}^{\prime},\, t_{j2}^{*}\equiv\vartriangle v_{i}v_{j}^{\prime}v_{j+1},\, t_{j3}^{*}\equiv\vartriangle v_{i}v_{j}v_{j+1}$
Triangulations of the quad with common vertex $v_{i}$ proposed by
{[}Xiong 2011{]} to define Mean LBO.}
\end{figure}

Applying the mean average area, according to \cite{Xiong2011}, from
all possible triangulations, as show in figure \ref{PAPER:fig:quad_xiong},
the area for quads $\ensuremath{A\left(Q_{v_{i}}\right)}$ and triangles
$A\left(T_{v_{i}}\right)$ is.

\noindent \begin{center}
$A\left(v_{i}\right)=\frac{1}{2^{m}}\overset{m}{\underset{j=1}{\sum}}2^{m-1}A\left(q_{j}\right)+\overset{r}{\underset{k=1}{\sum}}A\left(t_{k}\right)$
\par\end{center}

Where $q_{1},q_{2},...,q_{j},...,q_{m}\in Q_{v_{i}}$ and $t_{1},t_{2},...,t_{k},...,t_{r}\in T_{v_{i}}$

\begin{equation}
A\left(v_{i}\right)=\frac{1}{2}\overset{m}{\underset{j=1}{\sum}}\left[A\left(t_{j1}^{*}\right)+A\left(t_{j2}^{*}\right)+A\left(t_{j3}^{*}\right)\right]+\overset{r}{\underset{k=1}{\sum}}A\left(t_{k}\right)\label{PAPER:eq:area_1_ring_triangles_quads}
\end{equation}

Applying the gradient operator to (\ref{PAPER:eq:area_1_ring_triangles_quads})

\noindent \begin{center}
$\nabla A\left(v_{i}\right)=\frac{1}{2}\overset{m}{\underset{j=1}{\sum}}\left[\nabla A\left(t_{j1}^{*}\right)+\nabla A\left(t_{j2}^{*}\right)+\nabla A\left(t_{j3}^{*}\right)\right]+\overset{r}{\underset{k=1}{\sum}}\nabla A\left(t_{k}\right)$
\par\end{center}

According to (\ref{PAPER:eq:eq_gradient_voronoi_area}), we have

\noindent \begin{center}
$\nabla A\left(t_{j1}^{*}\right)=\frac{\cot\theta_{j3}\left(v_{j}-v_{i}\right)+\cot\theta_{j2}\left(v_{j}^{\prime}-v_{i}\right)}{2}$
\par\end{center}

\noindent \begin{center}
$\nabla A\left(t_{j2}^{*}\right)=\frac{\cot\theta_{j5}\left(v_{j}^{\prime}-v_{i}\right)+\cot\theta_{j4}\left(v_{j+1}-v_{i}\right)}{2}$
\par\end{center}

\noindent \begin{center}
$\nabla A\left(t_{j3}^{*}\right)=\frac{\cot\theta_{j6}\left(v_{j}-v_{i}\right)+\cot\theta_{j1}\left(v_{j+1}-v_{i}\right)}{2}$
\par\end{center}

\noindent \begin{center}
$\nabla A\left(t_{k}\right)=\frac{\cot\alpha_{k}\left(v_{k}-v_{i}\right)+\cot\beta_{k+1}\left(v_{k+1}-v_{i}\right)}{2}$
\par\end{center}

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/beltrami}

\protect\caption{\label{PAPER:fig:LBO-basic-5-TQ-1}The 5 basic triangle-quad cases
with common vertex $V_{i}$ and the relationship with $V_{j}$ and
$V_{j}^{\prime}$. (a) Two triangles {[}Desbrun 1999{]}. (b) (c) Two
quads and one quad {[}Xiong 2011{]}. (d) (e) Triangles and quads (TQLBO)
our contribution.}
\end{figure}

Triangle and quad configurations of the 1-ring neighborhood faces,
adjacent to $v_{i}$, can be simplified to five cases, as shown in
figure \ref{PAPER:fig:LBO-basic-5-TQ-1}.

According to equation (\ref{PAPER:eq:discrete_mean_curvature_normal}),
(\ref{PAPER:eq:def_LBO}), and five simple cases defined in figure
\ref{PAPER:fig:LBO-basic-5-TQ-1} the TQLBO (Triangle-Quad LBO) of
$v_{i}$ is

\noindent \begin{center}
$\triangle_{S}\left(v_{i}\right)=2\kappa\mathbf{n}=\frac{\nabla A}{A}=\underset{v_{j}\in N_{1}\left(v_{i}\right)}{\frac{1}{2A}\sum}w_{ij}\left(v_{j}-v_{i}\right)$
\par\end{center}

\begin{equation}
w_{ij}=\begin{cases}
\left(\cot\alpha_{j}+\cot\beta_{j}\right) & \mbox{case }\mathit{a.}\\
\frac{1}{2}\left(\cot\theta_{\left(j-1\right)1}+\cot\theta_{\left(j-1\right)4}+\cot\theta_{j3}+\cot\theta_{j6}\right) & \mbox{case \ensuremath{\mathit{b}}.}\\
\left(\cot\theta_{j2}+\cot\theta_{j5}\right) & \mbox{case \ensuremath{\mathit{c}}.}\\
\frac{1}{2}\left(\cot\theta_{j3}+\cot\theta_{j6}\right)+\cot\beta_{j} & \mbox{case \ensuremath{\mathit{d}}.}\\
\frac{1}{2}\left(\cot\theta_{\left(j-1\right)1}+\cot\theta_{\left(j-1\right)4}\right)+\cot\alpha_{j} & \mbox{case \ensuremath{\mathit{e}}.}
\end{cases}\label{PAPER:eq:TQLBO_wij}
\end{equation}

We define a TQLBO as a matrix equation

\begin{equation}
L\left(i,j\right)=\begin{cases}
-\frac{1}{2A_{i}}w_{ij} & \mbox{if }j\in N\left(v_{i}\right)\\
\frac{1}{2A_{i}}\underset{j\in N\left(v_{i}\right)}{\sum}w_{ij} & \mbox{if }i=j\\
0 & \mbox{otherwise}
\end{cases}\label{PAPER:eq:TQLBO_Simple_Matrix}
\end{equation}

Where $L$ is a $n\times n$ matrix, $n$ is the number of vertices
of a given mesh $M$, $w_{ij}$ is the TQLBO defined in equation (\ref{PAPER:eq:TQLBO_wij}),
$N\left(v_{i}\right)$ is the 1-ring neighborhood with shared face
to $v_{i}$, $A_{i}$ is the ring area around $v_{i}$.

Normalized equation of the TQLBO

\noindent 
\begin{equation}
L\left(i,j\right)=\begin{cases}
-\frac{w_{ij}}{\underset{j\in N\left(v_{i}\right)}{\sum}w_{ij}} & \mbox{if }j\in N\left(v_{i}\right)\\
\delta_{ij} & \mbox{otherwise}
\end{cases}\label{PAPER:eq:TQLBO-Normalized_Matrix}
\end{equation}

Where $\delta_{ij}$ being the Kronecker delta function.

\subsection{The Shape Inflation\label{PAPER:sub:Curvature-Enhancing}}

The shape is inflated by using the reverse direction of the curvature
flow, moving the vertices towards those mesh portions with larger
curvature. A standard diffusion process is applied:

\noindent \begin{center}
$\frac{\partial V}{\partial t}=\lambda L\left(V\right)$ 
\par\end{center}

To solve this equation, implicit integration is used as well as a
normalized version of TQLBO matrix

\begin{equation}
\left(I-\left|\lambda dt\right|W_{p}L\right)V^{\prime}=V^{t}\label{PAPER:eq:Lineal_System_with_wp}
\end{equation}

\noindent \begin{center}
$V^{t+1}=V^{t}+\mbox{sign}\left(\lambda\right)\left(V^{\prime}-V^{t}\right)$ 
\par\end{center}

The vertices $V^{t+1}$ are inflated, along their reverse curvature
direction, by solving the linear system: $Ax=b$, where $A=I-\left|\lambda dt\right|W_{p}L$,
$L$ is the Normalized TQLBO defined in the equation (\ref{PAPER:eq:TQLBO-Normalized_Matrix}),
$x=V^{\prime}$ are the smoothing vertices, $b=V^{t}$ are the actual
vertices positions, $W_{p}$ is a diagonal matrix with vertex weights,
and $\lambda dt$ is the inflate factor that supports negative and
positive values: negative for inflation and positive for smoothing.

The method was devised to use with weighted vertex groups, which specify
the final shape inflation of the solution, meaning \$0\$ as no changes
and 1 when a maximal change is applied. The weights modify the influence
zones, where the Laplacian is applied, as shown in equation \ref{PAPER:eq:Lineal_System_with_wp}
. Interestingly, the generated family of shapes may change substantially
with the weights of specific control points.

The curvature cannot be calculated at the boundary of the meshes that
are not closed, for that reason we use the scale-dependent operator
proposed by Desbrun et al. \cite{Desbrun1999}, the inflation factor
for boundary is represented by $\lambda_{e}$.

The model volume increases as the lambda is larger and negative, this
can be counteracted with a simple volume preservation. However, the
mesh may suffer large displacements when $\lambda<-1.0$ or after
multiple iterations. A simple volume conservation algorithm is: If
$v_{i}^{t+1}$ is a mesh vertex of $V^{t+1}$ in the $t+1$ iteration,
we define $\overline{v}$ as:

\noindent \begin{center}
$\overline{v}=\frac{1}{n}\underset{v_{i}\in V}{\sum}v_{i}$
\par\end{center}

$\overline{v}$is the mesh center, $vol_{ini}$ is an initial volume,
and $vol_{t+1}$ is the volume at the iteration $t+1$, $n$ is the
number of vertices, then the scale factor

\noindent \begin{center}
$\beta=\left(\frac{vol_{ini}}{vol_{t+1}}\right)^{\frac{1}{3}}$ 
\par\end{center}

allows to scale the vertices to:

\noindent \begin{center}
$v_{i\, new}^{t+1}=\beta\left(v_{i}^{t+1}-\overline{v}\right)+\overline{v}$ 
\par\end{center}

The shape inflation use a negative curvature flow that is an unstable
process when performing many iterations, however, our method uses
less than 3 iterations to get good results, and with 3 iterations
or less the method behaves stable.

\section{Sculpting\label{PAPER:sec:Sculpting}}

A new sculpting brush is herein proposed and aims to inflate the shape,
magnifying the shape curvatures of a polygon mesh in real time. This
brush works properly with the stroke method \textsl{Drag Dot}, allowing
to pre-visualize the model changes before the mouse is released. Also,
it allows to move the mouse along the model to match the shape zone
which is supposed to be inflated.

Brushes that perform a similar inflation can introduce mesh distortions
or produce mesh self-intersections, provided these brushes only move
the vertices along the normal without any global information. In contrast,
the present method searches for a proper inflation while preserving
the global curvature, retaining the original shape and main model
features. In addition, this method simplifies the work required for
the inflation since it needs not different brushes for inflating,
softening or styling. The inflated brush can make all these operation
in a single step. Real-time brushes require the Laplacian matrix is
constructed with the vertices that are within the sphere radius defined
by the user, reducing the matrix to be processed, the center of this
sphere depending on the place where the user clicks on the canvas
and the three-dimensional mesh placed where the click is projected.
Special handling is required for the boundary vertices with neighbors
that are not within the brush radius: these vertices are marked as
boundary and the curvature is not there calculated, but they must
be included in the matrix so that every vertex has their corresponding
neighbors within the selection. The sculpting Laplacian matrix reads
as.

\noindent \begin{center}
$\ensuremath{L\left(i,j\right)=\begin{cases}
-\frac{w_{ij}}{\underset{j\in N\left(v_{i}\right)}{\sum}w_{ij}} & \mbox{if }\left\Vert v_{i}-u\right\Vert <r\wedge\left\Vert v_{j}-u\right\Vert <r\\
0 & \mbox{if }\left\Vert v_{i}-u\right\Vert <r\wedge\left\Vert v_{j}-u\right\Vert \geq r\\
\delta_{ij} & \mbox{otherwise}
\end{cases}}$
\par\end{center}

Where $v_{j}\in N\left(v_{i}\right)$, $u$ is the sphere center of
radius $r$. The matrices should remove rows and columns of vertices
that are not within the radius.

\section{Subdivision surfaces\label{PAPER:sec:Subdivision-surfaces}}

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/teaser_cup}

\protect\caption{\label{PAPER:fig:Subdivision-Cups} Family of cups generated with
our method, from a coarse model (a), (c): the shape, obtained from
the Catmull-Clark Subdivision (b), (d), is inflated. Soft constraints,
over the coarse model, is drawn in red and blue (c).}
\end{figure}

The Catmull-Clark subdivision transformation is used to smooth a surface,
as the limit of a sequence of subdivision steps \cite{Stam1998}.
This process is governed by a B-spline curve \cite{Loop1987}, performing
a recursive subdivision transformation that refines the model into
a linear interpolation that approximates a smooth surface. The model
smoothness is automatically guaranteed \cite{DeRose1998}.

Catmull-Clark subdivision surface methods generate smooth and continuous
models from a coarse model and produce quick results because of the
simplicity of implementation. Nevertheless, changes to the global
curvature are hardly implantable. The Catmull-Clark subdivision surfaces
together with shape inflation can easily generate families of shapes
by changing a single parameter, allowing to handle a model with very
few vertices. In practice, this would allow an artist to choose a
model from a similar set of options that would meet his/her needs
without having to change each of the control vertices. Likewise, the
presented method allows the use of vertex weight paint over the control
points. The weights can be applied to a coarse model, followed by
a Catmull-Clark subdivision where weights are interpolated, producing
weights with smooth changes in the influence zones, as shown in figure
\ref{PAPER:fig:Subdivision-Cups}.c.

In equation \ref{PAPER:eq:Lineal_System_with_wp}, $W_{p}$ is a diagonal
matrix with weights corresponding to each vertex. Weights at each
vertex produce a different solution so that the matrix must be placed
in the diffusion equation since families that are generated may change
substantially with weighted of specific control points.

\section{Results\label{PAPER:sec:Results}}

The results of the shape inflation method with the extension of the
Laplace Beltrami operator for hybrid quad/triangle meshes with several
example models shown in figures \ref{PAPER:fig:Spectrum}, \ref{PAPER:fig:Subdivision-Cups},
\ref{PAPER:fig:Catmull_Clark}, \ref{PAPER:fig:TQLBO_test}, \ref{PAPER:fig:camello_enhanced},
\ref{PAPER:fig:Sculpt_Brush}, \ref{PAPER:fig:Performance-sculpt},
\ref{PAPER:fig:(a)Monkey}, \ref{PAPER:fig:Animated_Camell}. The
shape inflation was assessed with TQLBO method on a PC with AMD Quad-Core
Processor @ 2.40 GHz and 8 GB RAM.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[width=1\textwidth]{resources/figs/cruz_lambda4}
\par\end{centering}

\protect\caption{\label{PAPER:fig:Catmull_Clark}(a) Original Model, (b) Model with
Catmull-Clark Subdivision. Models with Laplacian smoothing: (c) and
(d). Models with a first Laplacian filtering $\lambda=60.0$, $\lambda_{e}=12.0$
and before applying shape inflation: (e) and (f). }
\end{figure}

Figure \ref{PAPER:fig:TQLBO_test} shows the results when applying
the Laplace Beltrami Operator TQLBO of equation (\ref{PAPER:eq:TQLBO_Simple_Matrix})
in a model with a simple subdivision. In column (c) the Laplacian
smoothing was applied to a model consisting of only quads. In column
(d) the model was converted to triangles and then the Laplacian smoothing
was applied. In column (e) the model was randomly converted from some
quads into triangles and then the Laplacian smoothing is applied,
showing similar results to those meshes composed only of triangles
or quads.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[height=0.5\textheight]{resources/figs/test_triangles_quads}
\par\end{centering}

\protect\caption{\label{PAPER:fig:TQLBO_test}(a) Original Model. (b) Simple subdivision.
(c), (d) (e) Laplacian smoothing with $\lambda=7$ and 2 iterations:
(c) for triangles, (d) for quads, (e) for triangles and quads random
chosen.}
\end{figure}

Methods using the Catmull-Clark subdivision surface and the inflation
allows to modify the curvature that is obtained with the process of
subdivision, as shown in figure \ref{PAPER:fig:Subdivision-Cups}.
This test used a coarse cup model, in which the subdivision was performed,
followed by a Laplacian smoothing and inflation. In figure \ref{PAPER:fig:Subdivision-Cups}.c,
\ref{PAPER:fig:Subdivision-Cups}.d shows also the use of weight vertex
groups over coarse models, with subdivision surfaces that allowed
to generate the weights for the new interpolated vertices. These new
weights were used for the inflation obtained on the 6 cups that are
at the right of the figure \ref{PAPER:fig:Subdivision-Cups}.d.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[height=0.5\textheight]{resources/figs/camello_enhanced2}
\par\end{centering}

\protect\caption{\label{PAPER:fig:camello_enhanced}Top row: Original camel model in
left. Shape inflation with $\lambda=-30.0$, $\lambda=-100.0$, $\lambda=-400.0$.
Bottom row: Shape inflation with weight vertex group, $\lambda=-50.0$
and 2 iterations for the legs, $\lambda=-200.0$ and 1 iteration for
the head and neck.}
\end{figure}

Laplacian smoothing applied with simple subdivision (see figure \ref{PAPER:fig:Catmull_Clark}.c.)
may produce similar results to those obtained with Catmull-Clark (see
figure \ref{PAPER:fig:Catmull_Clark}.b.), whose models are in average
equal triangles. The one obtained with the Laplacian smoothing is
shown in panel (c), (d) and those curvature modified versions are
in (e) and (f). As can be observed, different versions of the original
sketch can be obtained by parameterizing a single model value, a great
advantage of the presented method. Figure \ref{PAPER:fig:camello_enhanced}
shows the generation of different versions of a camel according to
the $\lambda$ parameter. In the top row, it is shown the shape inflation
results, as $\lambda$ becomes larger and negative, the resultant
shape is observed as if the model would inflate the more convex parts,
as shown in figure \ref{PAPER:fig:Spectrum}. The larger the $\lambda$
parameter the larger the model feature inflations. The bottom row
of figure \ref{PAPER:fig:camello_enhanced} shows the use of weighted
vertex groups, specifying which areas will be inflated. On the left,
the inflation of the camel legs produces an organic aspect, notice
that the border is not distorted and smooth.

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/camello_walk2}

\protect\caption{\label{PAPER:fig:Animated_Camell}The method is pose insensitive.
The inflation for the different poses are similar in terms of shape.
Top row: Original walk cycle camel model. Bottom row: Shape inflation
with weight vertex group, $\lambda=-400$ and $2$ iterations. }
\end{figure}

The inflation of the silhouette features is predictable and invariant
under isometric transformations, as those classically used in some
animations (see Figure \ref{PAPER:fig:Animated_Camell}). In this
figure, the animation shows some camel poses during a walk, the inflation
is performed at the neck and legs, as shown in the bottom left camel
in figure \ref{PAPER:fig:Animated_Camell}. Local modifications produced
by the pose interpolation or animation rigging practically do not
affect the result. In spite of at any pose of the camel legs there
is a clear difference, the inflation method allows a flesh-like shape
in the original pattern produced by the artist. This is due to the
mesh restricted diffusion process so that small local changes are
treated without affecting the global solution. The method therefore
is rotation invariant since it depends exclusively on the normal mesh
field, which is rotation invariant.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[height=0.3\textheight]{resources/figs/sculpt_brush}
\par\end{centering}

\protect\caption{\label{PAPER:fig:Sculpt_Brush}Top row: (a) Leg Camel, (b) Inflate
brush for leg into blue circle, (c) Inflate shape brush for leg into
red circle. Bottom row: (a) Hand, (b) Inflate brush for fingers into
blue circle, (c) Shape inflation brush for fingers in red circle.}
\end{figure}

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[width=1\columnwidth]{resources/figs/verts_per_second_sculpt}
\par\end{centering}

\protect\caption{\label{PAPER:fig:Performance-sculpt}Performance of our dynamic shape
inflation brush in terms of the sculpted vertices per second. Three
models with 12K, 40K, 164K vertices used for sculpting in real time.}
\end{figure}

Figure \ref{PAPER:fig:Sculpt_Brush} shows the use of a shape inflation
brush for sculpting in real time. One pass was used with the brush,
as shown in the figure, with the blue and red radius. In figure \ref{PAPER:fig:Sculpt_Brush}.b
the camel foot shows the inflation intersection that looks like two
bubbles, a similar pattern to what is observed to the fingers on the
bottom of the same figure. The silhouette inflation is observed in
figure \ref{PAPER:fig:Sculpt_Brush}.c since the main shape is retained
together with its finger and foot details. Similar results can be
obtained by a user, however it would take several steps and require
the use of several brushes, while the shape inflation took a single
step. Likewise, this new method can easily inflate organic features
like muscles during the sculpting process. In figure \ref{PAPER:fig:Performance-sculpt}
the shape inflation brush performance is illustrated, in this experiment
three models with 12K, 40K and 164K vertices, were used. These models
were sculpted with the shape inflate brush, at each step the brush
sphere containing a variable number of vertices for processing. The
processing time for 800 vertices in the camel paw (40k model) only
took 0.1 seconds, for 2600 vertices in the leg and neck (model 40k)
took 0.5 seconds, these times are suitable in real applications since
an artist sculpts a model for parts. \%, and each part is represented
by an average of about 1800 vertices.

\begin{figure}[H]
\includegraphics[width=1\columnwidth]{resources/figs/monkey}

\protect\caption{\label{PAPER:fig:(a)Monkey}(a) Bottom row: Original Model. Top row:
Original model scaled by 4. (b) Top and bottom row: inflating with
Normalized-TQLBO $\lambda=-50$ (c) Top and bottom row: inflating
with TQLBO $\lambda=-50$.}
\end{figure}

Tests with the Laplacian operator (equation \ref{PAPER:eq:TQLBO_Simple_Matrix})
and its normalized version (equation \ref{PAPER:eq:TQLBO-Normalized_Matrix}),
produce similar results if the triangles or quads that compose the
mesh are about the same size. The normalized version is more stable
and predictable because it is not divided by the area of the ring
which may be very small and causes numeric problems, as shown in figure
\ref{PAPER:fig:(a)Monkey}.c bottom row. The shape inflation of the
model with the normalized Laplacian operator results in a more regular
pattern. The model can be deformed with a TQLBO normalized version
with large $\lambda$ ($\lambda>400$) while intersecting itself with
no peaks. Figure \ref{PAPER:fig:(a)Monkey}.c shows different results
due to the quads areas in the model. Quads with larger area have smaller
inflation (figure \ref{PAPER:fig:(a)Monkey}.c skull), and smaller
quads have larger inflations (figure \ref{PAPER:fig:(a)Monkey}.c
chin).

\section{Implementation\label{PAPER:sec:Implementation}}

The method was implemented as a modifier for modeling and brush for
sculpting, on the Blender software \cite{blender} in C and C++. Working
with the Blender allowed to test the method interactively against
others, as Catmull-Clark, weight vertex groups and sculpting system
in Blender.

To improve the performance, it was worked with the Blender mesh structure,
visiting each triangle or quad and storing its corresponding index
and the sum of the Laplacian weights of the ring in a list so that
only two visits were required for the list of mesh faces and two times
the edge list, if the mesh was not closed. This drastically reduced
calculations, enabling real-time processing. In the construction of
the Laplacian matrix, several index were locked at vertices having
face areas or edge lengths with zero value that could cause spikes
and bad results.

Under these conditions, the matrix at equation \ref{PAPER:eq:TQLBO_Simple_Matrix}
is sparse since the number of neighbors per vertex, corresponding
to the number of data per row, is smaller compared to the total number
of vertices in the mesh. To solve the linear system equation \ref{PAPER:eq:Lineal_System_with_wp}
was used OpenNL \cite{Buatois2007} which is a a library for solving
sparse linear system.

\section{Conclusion and future work\label{PAPER:sec:Conclusion-and-future}}

This work presented an extension of the Laplace Beltrami operator
for hybrid quad/triangle meshes that can be used in production environments
and provides results similar to those obtained by working only with
triangles or quads. This paper has introduced a new way to change
silhouettes in a mesh for modeling or sculpting in a few steps by
means of the curvature model modification while preserving its overall
shape. In addition, a new modeling method has also been presented
some possible applications have been illustrated. The method works
properly with isometric transformations, opening the possibility of
introducing it on the process of animation.

We show that this tool may work in early modeling stages, case in
which coarse models are used, allowing to modify the shape generated
by the Catmull-Clark subdivision surfaces, and thereby avoiding edition
of the vertices with change of a single parameters.

Future work includes the analysis of theoretical relationships between
the Catmull-Clark subdivision surfaces and the Laplacian smoothing
since they can produce very similar results.

\section*{Acknowledgment}

We wold like to thank anonymous friends for their support of our research.

This work was supported in part by the Blender Foundation, Google
Summer of code program at 2012.

Livingstone elephant model is provided courtesy of INRIA and ISTI
by the AIM@SHAPE Shape Repository. Hand model is courtesy of the FarField
Technology Ltd. Camel model by Valera Ivanov is licensed under a Creative
Commons Attribution 3.0 Unported License. Dinosaur and Monkey models
are under public domain, courtesy of Blender Foundation.

\chapter{Mesh smoothing based on curvature flow operator in a diffusion equation\label{sub:Laplacian-Smooth}}

This work was accepted as part of the software Blender \cite{blender},
an open source 3D application for modeling, rendering, compositing,
video editing and game creation. The work was supported by an awarded
internship of the Google Summer of Code 2012 program, administered
for Google Inc.

\section{Synopsis}

Objects reconstructed from the real world contain undesirable noise
in many applications of computer graphics. A Mesh smoothing may remove
that noise while still preserving the geometry and shape of the original
model. This project aims to improve the mesh smoothing tools uses
by software Blender, using curvature flow operator in a diffusion
equation, allowing to work with hybrid meshes composed of triangles
and quads, using the Laplacian operator proposed by Pinzón and Romero
\cite{Pinzon2013}.

\section{Benefits to Blender}

This project proposes a new and robust mesh smoothing tool that improves
the appearance of surfaces models. Usually, methods to scan computer
graphics objects using the Kinect ZCam need to remove the noise present
at the time of capture. This mesh smoothing method produces higher
quality results without shrinkage, while the usually used smoothing
tool collapses after several iterations.

This mesh smoothing method allows hard and soft constraints on the
positions of the mesh points to maintain control over the shape, facilitating
to remove the noise generated during the sculpting, without eliminating
the desired details of the model.

\section{Deliverables }
\begin{itemize}
\item A new and robust mesh smoothing tool for Blender. 
\item Some documentation pages to be included in the manual. 
\item A technical document for developers to improve the method in the future. 
\item A tutorial explaining the use of the tool.
\end{itemize}

\section{Project Details }

The mesh smoothing algorithm was implemented as a diffusion equation
for specific geometric structures. The project was divided in four
parts:
\begin{enumerate}
\item Initialization of data and necessary structures.
\item Computation of the Laplacian Matrix.
\item Definition of the sparse linear system.
\item Solution of the sparse linear system, using a preconditioned bi-conjugated
gradient numerical library.
\end{enumerate}
Integration of the numerical library present in Blender to solve the
sparse linear system

Generation of documentation and tutorials.

\section{Project Schedule}
\begin{itemize}
\item 3 weeks: Understanding the Blender source code and identifying the
key points for the project.
\item 1 week: Definition of the data structures necessary to work with the
Blender architecture.
\item 1 week: Implementation of the methods for the initial configuration
of the smoothing algorithm. Implementation of the Laplacian matrix
calculation.
\item 2 week: Integration of the numerical library.
\item 2 week: formulation of the sparse linear system and implementation
of the numerical method to solve it.
\item 3 weeks: Formulation and implementation of the graphical user interface.
\item 2 weeks: Testing the tool.
\item 3 weeks: Generation of the documentation and tutorials.
\end{itemize}

\section{Mesh Smoothing}

A common way of attenuating noise in a polygonal mesh is by a diffusion
process \cite{Taubin1995,Desbrun1999}. Laplacian smooth techniques
allow a proper noise reduction of the mesh surface with minimal shape
changes. The simple idea consists in moving the vertices in the Laplacian
direction. When the cotangent version is used, the vertices are moved
in the direction of the curvature flow. The complexity of the Laplacian
smoothing can be linear in time and space with a fast convergence.
The diffusion process can attenuate noise with only one iteration
due the sparseness of the laplacian operator.

\noindent 
\begin{equation}
\frac{\partial V}{\partial t}=\lambda L\left(V\right)\label{eq:diffusion_process}
\end{equation}

Where $L$ is the Laplacian matrix defined in equation \ref{eq:LaplacianMatrixForSmoothing}
for meshes composed of triangles or quads with different sizes or
irregular sampling and $\lambda$ is a scalar that controls the diffusion
process, and smoothing factor. The equation \ref{eq:diffusion_process}
can be linearly approximated using implicit integration with a Laplacian
Operator version of TQLBO, the use of implicit integration makes the
system is more stable.

\begin{equation}
\left(I-\lambda dtL\right)V^{n+1}=V^{n}\label{eq:LaplacianSmoothLinearEquationSystem}
\end{equation}

The user may define the region of interest where the laplacian smooth
needs to be applied. For doing so, we add a diagonal matrix $W_{p}$
to equation \ref{eq:LaplacianSmoothLinearEquationSystem}, where every
element in the diagonal correspond to the weight for every vertex.

\noindent \begin{center}
$\left(I-\lambda dtW_{p}L\right)V^{n+1}=V^{n}$
\par\end{center}

For non-closed meshes or meshes with holes, the curvature flow can
not be computed. For this reason, the system smoothes out the edges
only in the direction of the diffusion process. The boundaries are
treated as a one-dimensional curve, where the Laplacian is defined
as the weighted difference between the vertex and the two immediate
neighbors, ensuring the curve maintains its original form as much
as possible. We define a Laplacian for mesh smoothing as a matrix
equation.

\noindent \begin{center}
\begin{equation}
L(i,j)=\begin{cases}
-\frac{1}{2A_{i}}w_{ij} & \mbox{if }j\in N(v_{i})\wedge v_{i}\notin\mbox{Boundary}\\
\frac{1}{2A_{i}}\underset{j\in N\left(v_{i}\right)}{\sum}w_{ij} & \mbox{if }i=j\wedge v_{i}\notin\mbox{Boundary}\\
-e_{ij} & \mbox{if }j\in N(v_{i})\wedge\left\{ v_{i},v_{j}\right\} \in\mbox{Boundary}\\
\frac{2}{E_{i}}\underset{j\in N\left(v_{i}\right)}{\sum}e_{ij} & \mbox{if }i=j\wedge\left\{ v_{i},v_{j}\right\} \in\mbox{Boundary}\\
0 & \mbox{otherwise}
\end{cases}\label{eq:LaplacianMatrixForSmoothing}
\end{equation}

\par\end{center}

Where $L$ is a $n\times n$ matrix, $n$ is the number of vertices
of a given mesh $M$, $w_{ij}$ is the TQLBO defined in equation (\ref{PAPER:eq:TQLBO_wij}),
$N\left(v_{i}\right)$ is the 1-ring neighborhood with shared face
to $v_{i}$, $e_{ij}=\frac{1}{\left\Vert v_{i}-v_{j}\right\Vert }$
is the inverse length of the edge between vertices $\left\{ v_{i},v_{j}\right\} $,
$E_{i}=\underset{j\in N\left(v_{i}\right)}{\sum}e_{ij}$. $A_{i}$
is the ring area around $v_{i}$.

\section{Results and Conclusions}

The developed user interface can be seen in figure \ref{fig:Interface-LaplacianSmoothing}.
This tool allows to set the $\lambda$ parameters for inner points
and boundaries, to configure soft constraints using weights defined
by vertices in ``Vertex Group'' and to set strong constraints by
independently applying the algorithm in the axis X, Y or Z.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics{resources/figs/Apinzonf_Diagram_Modifier_Panel}
\par\end{centering}

\protect\caption{\label{fig:Interface-LaplacianSmoothing}Panel inside blender user
interface of the Laplacian Smooth modifier tool.}
\end{figure}

The tool developed can set the $\lambda dt$ parameter of equation
\ref{eq:LaplacianSmoothLinearEquationSystem}. Using a small Lambda
factor ($\lambda<1.0$), noise can be removed without significantly
affecting the geometry (see figure \ref{fig:Factor-LaplacianSmoothing}.b),
while a large Lambda factor ($\lambda>1.0$) smoothed versions can
be obtained at the cost of losing fine geometry details (see figure
\ref{fig:Factor-LaplacianSmoothing}.c and \ref{fig:Factor-LaplacianSmoothing}.d).

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[width=1\textwidth]{resources/figs/LaplacianSmoothingFactor}
\par\end{centering}

\protect\caption{\label{fig:Factor-LaplacianSmoothing}Noise attenuation in face model
with Laplacian smoothing tool using only one iteration and changing
$\lambda$. (a) Original Model. (b) Smoothing $\lambda=0.5$. (c)
Smoothing $\lambda=2.5$ (d) Smoothing with $\lambda=5.0$. }
\end{figure}

The user can smooth the boundaries by configuring the parameter ``\textsl{Border}'',
seen in figure \ref{fig:Interface-LaplacianSmoothing}. Boundaries
are treated differently since there is no way to calculate the boundary
curvature flow. For this reason the Lambda factor ``\textsl{Border}''
just smooths them. The change of this parameter and the results seen
in figure \ref{fig:Border-LaplacianSmoothing}, illustrate how the
boundary inside the red circle is smoothed.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[width=1\textwidth]{resources/figs/LaplacianSmoothingBoundary}
\par\end{centering}

\protect\caption{\label{fig:Border-LaplacianSmoothing} Smoothing boundary changing
$\lambda_{Border}$ factor. (a) Original Model. (b) Smoothing $\lambda_{Border}=1.0$.
(c) Smoothing $\lambda_{Border}=2.5$ (d) Smoothing with $\lambda_{Border}=10.0$. }
\end{figure}

The tool allows the user to add soft constraints using weights for
each vertex, this allows to define regions of interest where you want
to apply the algorithm. In figure \ref{fig:VertexGroup-LaplacianSmoothing}.c,
the red region corresponds to the desired region and figure \ref{fig:VertexGroup-LaplacianSmoothing}.d
stands for those smoothed vertices in the red region.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[width=1\textwidth]{resources/figs/LaplacianSmoothingVertecGroup}
\par\end{centering}

\protect\caption{\label{fig:VertexGroup-LaplacianSmoothing}Use of weights per vertex
to constrain the effect of mesh smoothing. (a) Original Model. (b)
Smoothing with $\lambda=1.5$ (c) red vertices $weight=1.0$, blue
vertices $weight=0.0$. (d) Smoothing with $\lambda=2.5$. Red vertices
were only smoothing. }
\end{figure}

This module was developed as a tool for the Blender software to remove
noise in the most efficient way. The state-of-the-art methods remove
noise only if the mesh is composed by triangles, in contrast, with
the developed tool, artists can now remove the noise even if the mesh
is composed of triangles and quadrangles.

\chapter{Mesh Editing with Laplacian Deform\label{sub:Laplacian-Deform}}

This work was accepted and completed for the software Blender \cite{blender}
that is an open source 3D application for modeling, animation, rendering,
compositing, video editing and game creation. In the Google Summer
of Code 2013 program which was administered for Google Inc.

\section{Synopsis}

The mesh editing is generally done with affine transformations, Blender3D
offers some tools to transform the vertices as \textquotedblleft proportional
editing object mode\textquotedblright{} with which the transformation
of some vertices is interpolated to the other vertices connected with
the use of simple distance functions.

This project proposes to implement a method for mesh editing based
on sketching lines defines by user and preserving the geometric details
of the surface.

This method captures the geometric details using a differential coordinates
representations. The differential coordinates captures the local geometric
information (curvature and direction) of the vertex based on its neighbors.
This method allows you to retrieve the best possible original model
after changing the positions of some vertices using the differential
coordinates of the original model.

\section{Benefits to Blender}

This project proposes a new tool for blender user that requires preserve
the geometric details of the surface during a modeling, transformation,
definition of the shape keys of the mesh vertices.

The method will allow novice users to edit any polygon mesh preserving
the surface details.

This method allows the user to define new shape keys in a most fast
and intuitive way.

\section{Deliverables}
\begin{itemize}
\item A new mesh editing tool for Blender. 
\item Some pages of documentation to be included in the manual 
\item A technical document for developers to improve the method in the future. 
\item A tutorial explaining the use of the tool.
\end{itemize}

\section{Project Details}

The project would divide into several parts:
\begin{enumerate}
\item Calculate the differential coordinates. 
\item Store the fixed vertices (Hard constraints). 
\item Store positions of the edited vertices. 
\item Store the more representative vertex for retrieve rotation of every
differential coordinate.
\item Solve the initial solution \textendash{} in least-squares sense. 
\item Rotate the differential coordinates with base on initial solution
and more representative vertex.
\item Reconstruct the surface \textendash{} in least-squares sense. 
\item Generation of the documentation and tutorials.
\end{enumerate}

\section{Project Schedule}
\begin{itemize}
\item 2 Weeks: Calculate the differential coordinates.
\item 2 Weeks: Store the fixed vertexes (Hard constraints).
\item 2 Weeks: Store positions of the edited vertexes.
\item 2 Weeks: Compute initial solution.
\item 2 Weeks: Rotate differential coordinates.
\item 2 Weeks: Reconstruct the surface \textendash{} in least-squares sense. 
\item 1 Weeks: Testing the tool and Define and implement graphical user
integration.
\item 2 Weeks: Generation of the documentation and tutorials.
\end{itemize}

\section{Laplacian Deform}

The Laplacian deformation facilitates mesh visualization while it
preserves the geometric surface details. This method allows to define
a set of anchor vertices which serve as a repair to move the mesh
by translating some of them. The system keeps the anchor vertices
in fixed positions and calculates the best possible locations of the
other vertices for the shape to preserve the original geometric details.
This work adapts the method proposed by Sorkine\cite{Sorkine2004}
for mesh deformations at deleting the use of static vertices. The
method has been in addition applied to hybrid meshes composed by triangles
and quads, using the proposed TQLBO. In particular, the geometric
details are captured using differential coordinate representations
that contain the local geometric information (curvature and direction)
of the vertex and its neighbors, as shown in figure \ref{fig:DifferentialCoor}.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[width=0.7\columnwidth]{resources/figs/DifferentialCoordinates}
\par\end{centering}

\protect\caption{\label{fig:DifferentialCoor} Difference between $v_{i}$ and the
center of mass of its neighbors $v_{1},...,v$.}
\end{figure}

\begin{equation}
\delta_{i}=\overset{m}{\underset{j=1}{\sum}}w_{ij}\left(v_{i}-v_{j}\right)\label{eq:DifferentialCoordinate}
\end{equation}

Where $\delta_{i}$ is the differential coordinate for vertex $v_{i}$.
The $v_{j}$ are the immediate neighbors of $v_{i}$, and $w_{ij}$
is the weight between vertex $v_{i}$ and $v_{j}$ defined in equation
\ref{PAPER:eq:TQLBO_wij} that is TQLBO.

Then the linear system for finding the new pose of a mesh is.

\begin{equation}
\begin{bmatrix}w_{l}L\\
W_{c}
\end{bmatrix}X=\begin{bmatrix}\delta\\
W_{c}C
\end{bmatrix}\label{eq:LaplacianDefomSystem}
\end{equation}

Where $w_{l}$ is the Laplacian Matrix weight $L$, and the Laplacian
matrix $L$ was defined in equation \ref{PAPER:eq:TQLBO_Simple_Matrix}
. $Wc$ is a matrix that contains ones int the indexes of anchors
vertices. $C$ is a vector with coordinates of anchors vertices after
several transformations. $\delta$ are the differential coordinates
defined in equation \ref{eq:DifferentialCoordinate}.

\section{Testing Solvers }

For this project we chose a numerical solver to be included in the
Blender software after an evaluation of the initial factorization
of the Laplacian deformation system.

Linear equation system to solve

\noindent \begin{center}
$\begin{bmatrix}w_{l}L\\
W_{c}
\end{bmatrix}X=\begin{bmatrix}\delta\\
W_{c}C
\end{bmatrix}$
\par\end{center}

Solving the sparse linear system

\noindent \begin{center}
$Ax=b$
\par\end{center}

Where:

\noindent \begin{center}
$A=\begin{bmatrix}w_{l}L\\
W_{c}
\end{bmatrix}$
\par\end{center}

\noindent \begin{center}
$x=V$
\par\end{center}

\noindent \begin{center}
$b=\begin{bmatrix}\delta\\
W_{c}C
\end{bmatrix}$
\par\end{center}

\subsection{Hardware Specification }
\begin{itemize}
\item Processor: AMD Quad-Core 2.40 GHz 
\item RAM: 8.0 GB 
\item OS: Windows 7 Professional 
\item Graphics controller: NVIDIA Quadro FX 570
\end{itemize}

\subsection{Software Specification }
\begin{description}
\item [{CGAL}] Computational Geometry Algorithms Library 
\item [{Graphite}] Research platform for computer graphics
\end{description}

\subsection{Numeric Solvers Used }
\begin{description}
\item [{CG:}] Conjugate gradient method. 
\item [{BICGSTAB:}] Biconjugate gradient stabilized method. 
\item [{GMRES:}] Generalized minimal residual method. 
\item [{SUPERLU:}] Sparse Direct Solver, LU decomposition with partial
pivoting. 
\item [{TAUCS\_LDLT:}] A library of sparse linear solvers with LDLT factorization. 
\item [{CHOLMOD:}] Supernodal sparse Cholesky factorization. 
\end{description}
LU factorization is a numerical method that works with large, sparse,
non-symmetric systems of linear equations \cite{Levy2005}. We choose
the implementation of LU factorization in a OpenNL-SuperLU library,
because this method show the better performance for the computation
of a solution for a Laplacian Deform linear system of equations presented
in equation \ref{eq:LaplacianDefomSystem} how see in the table and
plot \ref{tab:TimeVsSolvers}, \ref{fig:Plot-IniFactor}. OpenNL SuperLU
allow works with the Graphics Unit Processor GPU, for exploit the
capacity of GPU to work with parallel structures, more fast that traditional
CPU.

\begin{table}[H]
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline 
Model & Vertices & CG & BICGSTAB & GMRES & SUPERLU & TAUCS & CHOLMOD\tabularnewline
\hline 
\hline 
Cross & 24 & 0.05 & 0.05 & 0.04 & 0.04 & 0.05 & 0.05\tabularnewline
\hline 
King & 538 & 0.83 & 0.63 & 0.71 & 0.61 & 0.68 & 0.79\tabularnewline
\hline 
YModel & 4770 & 19.60 & 16.44 & 16.93 & 16.06 & 16.88 & 17.95\tabularnewline
\hline 
Man & 10002 & 33.43 & 27.76 & 29.91 & 28.54 & 29.53 & 30.80\tabularnewline
\hline 
Neptune & 28052 & 133.97 & 136.46 & 136.39 & 133.21 & 142.87 & 142.76\tabularnewline
\hline 
Armadillo & 34594 & 194.48 & 174.88 & 175.80 & 169.92 & 181.70 & 183.49\tabularnewline
\hline 
\end{tabular}

\protect\caption{\label{tab:TimeVsSolvers}Vertices Vs Seconds, Laplacian Deform initial
factorization performance.}
\end{table}

\begin{figure}[H]
\includegraphics[width=1\textwidth]{Data/Benchmark}

\protect\caption{\label{fig:Plot-IniFactor}Plot of Vertices Vs Seconds, Initial factorization
performance.}
\end{figure}

\section{Results}

The user interface for the software Blender can be seen in figure
\eqref{fig:Panel-LaplacianDeform}\}. In this tool, the user defines
the anchor vertices and, with the use of a vertex, groups the features
offered by Blender. The user configures this name in the field \textsl{Anchors
Vertex Group}. The \textsl{Bind} option initiates the system and captures
the geometry details in form of differential coordinates and computes
the factorization of the linear system, after that the system is ready
to pose meshes in a real-time interaction session.

\noindent \begin{center}
\begin{figure}[H]
\noindent \begin{centering}
\includegraphics{resources/figs/Apinzonf_Diagram_Deform_Modifier_Panel_01}
\par\end{centering}

\protect\caption{\label{fig:Panel-LaplacianDeform}Panel inside blender user interface
of the Laplacian Deform modifier tool.}
\end{figure}

\par\end{center}

Figure \eqref{fig:Armadillo} shows the Laplacian Deformation applied
to a model with 173K vertices, only anchor vertices were used and
are represented in blue color. When the user applies several transformations
(location, rotation, scale) to this anchors vertices, the system finds
a solution and estimates the position of the vertices (in yellow).
This method works in real time and for doing so the matrix $\begin{bmatrix}w_{l}L\\
W_{c}
\end{bmatrix}$ in equation \eqref{eq:LaplacianDefomSystem} is $LU$ decomposed
only once, with LU factorization,, when the system initiates. Once
the matrix is factorized, the system can solve the unknowns in the
order of milliseconds. Results are improved by solving several times
the system of equations, without LU factorization at every iteration,
just the differential coordinates are adjusted since the differential
coordinates can be rotated. Only four iterations were necessary for
obtaining figure \eqref{fig:Armadillo} but the system finds proper
solutions with only one iteration, when the angle of rotations are
less that$\pi$.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[width=1\textwidth]{resources/figs/Armadillo}
\par\end{centering}

\protect\caption{\label{fig:Armadillo}Anchor vertices in blue color. (a) Original
Model, (b,c,d) new poses change only anchor vertices, system find
positions for vertices in yellow color.}
\end{figure}

Figure \ref{fig:Cactus} shows a comparison after making a single
transformation that consists in rotating $70\lyxmathsym{º}$ to the
right the parts, in blue color. Results using a bicubic interpolation
are shown in \ref{fig:Cactus}.b observe how the main trunk losses
its shape when rotated. The propagation of changes made by the Laplacian
deformation (figure \ref{fig:Cactus}.c) for the same transformation,
shows better results.

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/cactus}

\protect\caption{\label{fig:Cactus}(a) Original cactus model. (b) Rotate 70º to right
the blue segments with basic interpolation (c) Rotate 70º to right
the blue segments with Laplacian deform tool.}
\end{figure}

The Laplacian Deformation tool allows an user to pose a model while
preserve the geometry details, in figure \ref{fig:Horse}.b y \ref{fig:Horse}.c
Observe the horse new pose after five transformations and rotation
of the head, in figure \ref{fig:Horse}.b using basic interpolation,
the shape and details are lost at every change. In constrast, figure
\ref{fig:Horse}..c illustrates how the new pose of the horse looks
more natural, above all for the body and neck. This comparison indicates
this method allows to apply any transformation without loss of details.

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/horse}

\protect\caption{\label{fig:Horse}(a) Original Horse model. (b) Translate and rotate
the blue segments with basic interpolation (c) Translate and rotate
the blue segments with Laplacian deform tool.}
\end{figure}

\chapter{Skeleton Extraction}

Part of this work was accepted and presented as a poster titled\textsl{
Software para la Extracción del Esqueleto por Contracción y Suavizado}
at 7th International Seminar on Medical Image Processing and Analysis
SIPAIM 2011. The conference was held on December 6-8 of 2011, in Bucaramanga,
Colombia \cite{Pinzon2011}, the poster thumbnail is shown in figure
\ref{fig:PosterSIPAIM2011}.

\begin{figure}
\noindent \begin{centering}
\includegraphics[width=0.9\textwidth]{poster/SIPAIM2011/PosterSIPAIM2011}
\par\end{centering}

\protect\caption{\label{fig:PosterSIPAIM2011}Poster \textsl{Software para la Extracción
del Esqueleto por Contracción y Suavizado} presented at 7th International
Seminar on Medical Image Processing and Analysis SIPAIM 2011.}

\end{figure}

Part of this work was accepted and presented as a poster titled\textsl{
Análisis Experimental de la Extracción del Esqueleto por Contracción
con Suavizado Laplaciano} at 6th International Seminar on Medical
Image Processing and Analysis SIPAIM 2010. The conference was held
on December 1-4 of 2011, in the city of Bogotá, Colombia \cite{Pinzon2010},
the poster thumbnail is shown in figure \ref{fig:PosterSIPAIM2010}.

\begin{figure}
\noindent \begin{centering}
\includegraphics[width=0.9\textwidth]{poster/SIPAIM2010/FINAL/PosteraA0-150dpi}
\par\end{centering}

\protect\caption{\label{fig:PosterSIPAIM2010}Poster \textsl{Análisis Experimental
de la Extracción del Esqueleto por Contracción con Suavizado Laplaciano}
presented at 6th International Seminar on Medical Image Processing
and Analysis SIPAIM 2010.}
\end{figure}

\section{Background}

Skeleton extraction not only reduces the dimensionality but rather
it represents a three-dimensional object as a uni-dimensional structure
\cite{Cornea2007}.

The skeleton extraction use the natural shrink produced by the laplacian
smoothing \cite{Meyer2003}, to iteratively contract the mesh until
the volume inside the surface is about zero (see figure \ref{fig:MeshContracted}).
The method stretches the mesh to warranty that object preserves, as
much as possible, the original topology constraint \cite{Au2008}.

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/contraccion}

\protect\caption{\label{fig:MeshContracted}From left to right iterative mesh contraction.}
\end{figure}

The laplacian smooth process moves the vertices in the direction of
the Laplacian, if the cotangent laplacian operator is used, the vertices
move in the direction of minimal surface \cite{Pinkall1993} following
the curvature flow of the mesh surface.

In \cite{Au2008} it is proposed the next system of equations to iteratively
contract the mesh until a skeleton.

\begin{equation}
\left[\begin{array}{c}
W_{L}L\\
W_{H}
\end{array}\right]X_{t+1}=\left[\begin{array}{c}
0\\
W_{H}X_{t}
\end{array}\right]\label{eq:SkeletonExtraction}
\end{equation}

where $L$ is the Laplacian matrix describe in equation \ref{PAPER:eq:TQLBO_Simple_Matrix},
$W_{L}$ is the smoothing factor, $W_{H}$ is the attraction constraint
factor.

$W_{L}$ and $W_{H}$ are changing at every iteration. The contraction
and smoothing constraint $W_{L}^{t+1}=S_{L}W_{L}^{t}$ with $S_{L}=2.0$.
The attraction constraint $W_{H,i}^{t+1}=W_{H,i}^{0}\sqrt{\frac{A_{i}^{0}}{A_{i}^{t}}}$.

$A_{i}^{t}$ y $A_{i}^{0}$ are the current area and initial area
of the ring surrounding $x_{i}$.

\section{Contribution}

This work proposed an additional constraint which seeks the vertices
while the skeleton does not shrink. Such constraint eliminates the
need of adjusting the final skeleton, taking each skeleton node and
move it to the center of the local mesh region.

The basic idea is move the vertices along of the normal line, estimated
at every vertex, based on the average of the normals of the faces.

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/LineConstraintSkeleton}

\protect\caption{\label{fig:LineConstraint}Left: The vertex $x_{i}$ move along line
constraint. Right: the distance of vertex $x_{i}$ to plane 1 and
plane 2 when change the position in every iteration.}
\end{figure}

This line can be see as intersection of two planes (see figure \ref{fig:LineConstraint})
created with the information of the vertex $x_{i}$, the normal $\overrightarrow{\mathbf{n}_{i}}$
and neighbors $x_{j}$ of $x_{i}$.

The plane equation.

\noindent \begin{center}
$a\mathbf{x}+b_{0}\mathbf{y}+c_{0}\mathbf{z}+d_{0}=0$
\par\end{center}

There are various methods that build a plane equation using three
non-collinear points. These three points were herein chosen using
the vertex $x_{i}$, the normal $\overrightarrow{\mathbf{n}_{i}}$
and neighbors $x_{j}$ of $x_{i}$.

$P_{1}=x_{i}$, $P_{2}=x_{j}$ when $x_{j}\in Neighboors\left(x_{i}\right)$
and $P_{3}=P_{1}-\overrightarrow{\mathbf{n}_{i}}$.

A equation plane only requires three non-collinear points$\left\{ P_{1},P_{2},P_{3}\right\} $
and solve the next systems of equations for the $a,b,c$ variables.

\noindent \begin{center}
$\begin{array}{c}
ax_{1}+by_{1}+cz_{1}+d=0\\
ax_{2}+by_{2}+cz_{2}+d=0\\
ax_{3}+by_{3}+cz_{3}+d=0
\end{array}$
\par\end{center}

The distance of point $P_{0}=\{x_{0},y_{0},z_{0}\}$ to a plane $\Pi=a\mathbf{x}+b_{0}\mathbf{y}+c_{0}\mathbf{z}+d_{0}$.

\noindent \begin{center}
$\left|\Pi-P_{0}\right|=\frac{\left|ax_{0}+by_{0}+cz_{0}+d\right|}{\sqrt{a^{2}+b^{2}+c^{2}}}$
\par\end{center}

In the equation above, every point $P_{0}=\{x_{0},y_{0},z_{0}\}$
that belongs to the plane defined when this equation is null. If the
point $P_{0}$ is not in the plane, the value of the plane equation
changes, and the absolute value is increased if $P_{0}$ is farther
with respect to the plane. Let us use this simple relation to build
a valid constraint that can be put in the form of the linear equation
$Ax=B$. Where $A$ is the matrix with values $\begin{vmatrix}a & b & c\end{vmatrix}$
the $x$ value correspond to every vertex in mesh $\{x_{0},y_{0},z_{0}\}$
and $B$ stands for the $d$ value of the plane equation. Provided
that the linear system presented in equation \ref{eq:SkeletonExtraction}
has many solutions, the system solves it in the least-squares sense.
With the new constraint, the system then tries to minimize the distance
between the new vertex positions and the two planes.

The new system of equations proposed

\begin{equation}
\left[\begin{array}{c}
W_{L}L\\
W_{H}\\
W_{D}\Pi_{1}\\
W_{D}\Pi_{2}
\end{array}\right]X_{t+1}=\left[\begin{array}{c}
0\\
W_{H}X_{t}\\
-D_{1}\\
-D_{2}
\end{array}\right]\label{eq:SkeletonExtractionDistance}
\end{equation}

where $W_{D}$ is the weight to new constraints, that forced the vertices
to move along of the normal line of every vertex. $\Pi_{1}$ and $\Pi_{2}$
are matrix that contains $a,b,c$ values of the equation of the plane
for every vertex. $D_{1}$ and $D_{2}$ are the vector with $d$ values
of the equations of the planes for every vertex.

\section{Results and Conclusions}

\begin{figure}
\includegraphics[width=1\textwidth]{resources/figs/Skeletonizer_1}

\protect\caption{\label{fig:Skeletonizer-interface}sjSkeletonizer interface. (a) Open
and display mesh. (b) Mesh contracted after 4 iterations. (c) Skeleton
get after surface simplification.}
\end{figure}

As a result of this work, the software\textsl{ sjSkeletonizer} \cite{Pinzon2012}
permits the processing, visualization and extraction of the skeleton
from polygonal hybrid meshes composed of triangles and quads. In figure
\ref{fig:Skeletonizer-interface}, the software interface of can be
observed. The software Skeletonizer allows to open and display meshes
in an interactive viewer. This software contracts a mesh with the
use of two different Laplacian operators: The cotangent operator and
umbrella operator, the user can customize the number of iterations
of the contraction step. The software allows an user to visualize
the original mesh and contract mesh after a particular number of iterations
(figure \ref{fig:Skeletonizer-interface}.b). In the final step of
the process, the software simplifies the mesh (figure \ref{fig:Skeletonizer-interface}.c)
using the same method proposed by \cite{Au2008} that consist in a
modification version of quadric error metrics of Garland \cite{Garland1997}
.

\begin{figure}[H]
\includegraphics[width=1\textwidth]{resources/figs/poses}

\protect\caption{\label{fig:Skeleton-extraction-poses} Model with different poses
and skeleton obtained with our skeleton extraction software.}
\end{figure}

Figure \ref{fig:Skeleton-extractedModels} shows the skeleton extracted
from different models using our method. Every skeleton accurately
represents the topology and geometry of the original meshes. The extracted
skeletons are homotopic to the original shapes. Our method permits
also to simplify the skeleton until the user is required to guarantee
that every node in the skeleton represents as much as possible every
core part of the original model.

\begin{figure}
\includegraphics[width=1\textwidth]{resources/figs/Skeletonizer_2}

\protect\caption{\label{fig:Skeleton-extractedModels}Skeleton extracted from different
models. (a) Dog model (b) Character model. (c) Person model. (d) Clay
model.}
\end{figure}

The method proposed is robust to isometric transformations ( figure
\ref{fig:Skeleton-extraction-poses}) so that if some global transformations
like rotations of central parts of the mesh are made, the extracted
skeletons are similar. probably because of the properties of the Laplacian
operator. The figure shows similar skeletons for different poses of
the same model, these skeletons show identical branching structure.

\chapter{Conclusion \label{sec:Conclusion-and-future}}

This work presented a novel extension of the Laplace Beltrami operator
for hybrid quad/triangle meshes TQLBO, and the successful applications
of such principle in different type of problems in computer geometric
modelling like smoothing, enhancing, sculpting, deformation, reposing
and skeleton extraction.

We have largely demonstrated a good performance, stability and robustness
of the extension proposed. This novel extension of the Laplace Beltrami
operator was introduced in the computer modeling industry inside the
3D computer graphics software Blender, thereby facilitating artists
to work with different tools developed using TQLBO, for any kind of
meshes, i.e., triangular meshes, quadrilateral meshes or hybrid meshes.

\bibliographystyle{plain}
\addcontentsline{toc}{chapter}{\bibname}\bibliography{bibliografia/listado_inicial,TQLBO-paper/template,D:/src/blender/GSoC/GSOC2013/bibliography_mesh_editing,SIBGRAPI2013/template}

\end{document}